### 2018-09-14 proposal/proposal.tex
Initial commit

% TEMPLATE for Usenix papers, specifically to meet requirements of
%  USENIX '05
% originally a template for producing IEEE-format articles using LaTeX.
%   written by Matthew Ward, CS Department, Worcester Polytechnic Institute.
% adapted by David Beazley for his excellent SWIG paper in Proceedings,
%   Tcl 96
% turned into a smartass generic template by De Clarke, with thanks to
%   both the above pioneers
% use at your own risk.  Complaints to /dev/null.
% make it two column with no page numbering, default is 10 point

% Munged by Fred Douglis <douglis@research.att.com> 10/97 to separate
% the .sty file from the LaTeX source template, so that people can
% more easily include the .sty file into an existing document.  Also
% changed to more closely follow the style guidelines as represented
% by the Word sample file.

% Note that since 2010, USENIX does not require endnotes. If you want
% foot of page notes, don't include the endnotes package in the 
% usepackage command, below.

% This version uses the latex2e styles, not the very ancient 2.09 stuff.
\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{usenix,epsfig,endnotes}
\begin{document}

%don't want date printed
\date{}

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{\Large \bf Wonderful : A Terrific Application and Fascinating Paper}

%for single author (just remove % characters)
\author{
{\rm Your N.\ Here}\\
Your Institution
\and
{\rm Second Name}\\
Second Institution
}

\maketitle

% Use the following at camera-ready time to suppress page numbers.
% Comment it out when you first submit the paper for review.
% \thispagestyle{empty}

\subsection*{Abstract}
Your Abstract Text Goes Here.  Just a few facts.
Whet our appetites.

\section{Introduction}

A paragraph of text goes here.  Lots of text.  Plenty of interesting
text. \\

More fascinating text. Features\endnote{Remember to use endnotes, not footnotes!} galore, plethora of promises.\\

\section{This is Another Section}

Some embedded literal typset code might 
look like the following :

{\tt \small
\begin{verbatim}
int wrap_fact(ClientData clientData,
              Tcl_Interp *interp,
              int argc, char *argv[]) {
    int result;
    int arg0;
    if (argc != 2) {
        interp->result = "wrong # args";
        return TCL_ERROR;
    }
    arg0 = atoi(argv[1]);
    result = fact(arg0);
    sprintf(interp->result,"%d",result);
    return TCL_OK;
}
\end{verbatim}
}

Now we're going to cite somebody.  Watch for the cite tag.
Here it comes~\cite{Chaum1981,Diffie1976}.  The tilde character (\~{})
in the source means a non-breaking space.  This way, your reference will
always be attached to the word that preceded it, instead of going to the
next line.

\section{This Section has SubSections}
\subsection{First SubSection}

Here's a typical figure reference.  The figure is centered at the
top of the column.  It's scaled.  It's explicitly placed.  You'll
have to tweak the numbers to get what you want.\\

% you can also use the wonderful epsfig package...
\begin{figure}[t]
\begin{center}
\begin{picture}(300,150)(0,200)
\put(-15,-30){\special{psfile = fig1.ps hscale = 50 vscale = 50}}
\end{picture}\\
\end{center}
\caption{Wonderful Flowchart}
\end{figure}

This text came after the figure, so we'll casually refer to Figure 1
as we go on our merry way.

\subsection{New Subsection}

It can get tricky typesetting Tcl and C code in LaTeX because they share
a lot of mystical feelings about certain magic characters.  You
will have to do a lot of escaping to typeset curly braces and percent
signs, for example, like this:
``The {\tt \%module} directive
sets the name of the initialization function.  This is optional, but is
recommended if building a Tcl 7.5 module.
Everything inside the {\tt \%\{, \%\}}
block is copied directly into the output. allowing the inclusion of
header files and additional C code." \\

Sometimes you want to really call attention to a piece of text.  You
can center it in the column like this:
\begin{center}
{\tt \_1008e614\_Vector\_p}
\end{center}
and people will really notice it.\\

\noindent
The noindent at the start of this paragraph makes it clear that it's
a continuation of the preceding text, not a new para in its own right.

Now this is an ingenious way to get a forced space.
{\tt Real~$*$} and {\tt double~$*$} are equivalent.

Now here is another way to call attention to a line of code, but instead
of centering it, we noindent and bold it.\\

\noindent
{\bf \tt size\_t : fread ptr size nobj stream } \\

And here we have made an indented para like a definition tag (dt)
in HTML.  You don't need a surrounding list macro pair.
\begin{itemize}
\item[]  {\tt fread} reads from {\tt stream} into the array {\tt ptr} at
most {\tt nobj} objects of size {\tt size}.   {\tt fread} returns
the number of objects read. 
\end{itemize}
This concludes the definitions tag.

\subsection{How to Build Your Paper}

You have to run {\tt latex} once to prepare your references for
munging.  Then run {\tt bibtex} to build your bibliography metadata.
Then run {\tt latex} twice to ensure all references have been resolved.
If your source file is called {\tt usenixTemplate.tex} and your {\tt
  bibtex} file is called {\tt usenixTemplate.bib}, here's what you do:
{\tt \small
\begin{verbatim}
latex usenixTemplate
bibtex usenixTemplate
latex usenixTemplate
latex usenixTemplate
\end{verbatim}
}

\subsection{Last SubSection}

Well, it's getting boring isn't it.  This is the last subsection
before we wrap it up.

\section{Acknowledgments}

A polite author always includes acknowledgments.  Thank everyone,
especially those who funded the work.

\section{Availability}

It's great when this section says that MyWonderfulApp is free software, 
available via anonymous FTP from

\begin{center}
{\tt ftp.site.dom/pub/myname/Wonderful}\\
\end{center}

Also, it's even greater when you can write that information is also 
available on the Wonderful homepage at

\begin{center}
{\tt http://www.site.dom/\~{}myname/SWIG}
\end{center}

Now we get serious and fill in those references.  Remember you will
have to run latex twice on the document in order to resolve those
cite tags you met earlier.  This is where they get resolved.
We've preserved some real ones in addition to the template-speak.
After the bibliography you are DONE.

{\footnotesize \bibliographystyle{acm}
\bibliography{../common/bibliography}}

\theendnotes

\end{document}

### 2018-09-15 proposal/proposal.tex
Put in all information we have so far

@@
 \usepackage{usenix,epsfig,endnotes}
+\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}
 \begin{document}
@@
 %make title bold and 14 pt font (Latex default is non-bold, 16 pt)
-\title{\Large \bf Wonderful : A Terrific Application and Fascinating Paper}
+\title{\Large \bf CS380S: Project Proposal}
 
@@
 \author{
-{\rm Your N.\ Here}\\
-Your Institution
+{\rm Rushi Shah}
 \and
-{\rm Second Name}\\
-Second Institution
+{\rm Andrew Russell}
 }
@@
 
-% Use the following at camera-ready time to suppress page numbers.
-% Comment it out when you first submit the paper for review.
-% \thispagestyle{empty}
+\section{Project Idea}
 
-\subsection*{Abstract}
-Your Abstract Text Goes Here.  Just a few facts.
-Whet our appetites.
+We plan to use Data Dependency tools to determine how system entropy in a given program is used by various cryptographic algorithms.
+That will allow us to identify if and when entropy is too low or is misused.
+We can release this as a tool for developers to use as part of a compiler toolchain.
+We can also use the tool on current projects that might be misusing crypto/entropy.
 
-\section{Introduction}
+We have two motivating examples.
+First, we would like our tool to be able to detect the issue with the Debian/OpenSSL pseudo-random number generator that was exposed in 2008.
+Second, we would like to identify potential vulnerabilities in current cryptocurrency wallet code.
 
-A paragraph of text goes here.  Lots of text.  Plenty of interesting
-text. \\
+\section{Rough Plan}
 
-More fascinating text. Features\endnote{Remember to use endnotes, not footnotes!} galore, plethora of promises.\\
+% TODO
 
-\section{This is Another Section}
+\section{Research Hypothesis}
 
-Some embedded literal typset code might 
-look like the following :
+% TODO
 
-{\tt \small
-\begin{verbatim}
-int wrap_fact(ClientData clientData,
-              Tcl_Interp *interp,
-              int argc, char *argv[]) {
-    int result;
-    int arg0;
-    if (argc != 2) {
-        interp->result = "wrong # args";
-        return TCL_ERROR;
-    }
-    arg0 = atoi(argv[1]);
-    result = fact(arg0);
-    sprintf(interp->result,"%d",result);
-    return TCL_OK;
-}
-\end{verbatim}
-}
+\section{Related Work}
 
-Now we're going to cite somebody.  Watch for the cite tag.
-Here it comes~\cite{Chaum1981,Diffie1976}.  The tilde character (\~{})
-in the source means a non-breaking space.  This way, your reference will
-always be attached to the word that preceded it, instead of going to the
-next line.
+% TODO maybe we need to say what exactly we found in the links instead of just dropping links?
+\subsection{Background information}
 
-\section{This Section has SubSections}
-\subsection{First SubSection}
+\begin{enumerate}
+	\item Debian/OpenSSL Bug 
+		\begin{enumerate}
+			\item \url{https://www.schneier.com/blog/archives/2008/05/random_number_b.html}
+			\item \url{https://research.swtch.com/openssl}
+			\item \url{https://freedom-to-tinker.com/2013/09/20/software-transparency-debian-openssl-bug/}
+			\item \url{https://www.cs.umd.edu/class/fall2017/cmsc818O/papers/private-keys-public.pdf}
+		\end{enumerate}
+	\item Data flow
+		\begin{enumerate}
+			\item \url{https://en.wikipedia.org/wiki/Data-flow_analysis}
+			\item \url{https://www.seas.harvard.edu/courses/cs252/2011sp/slides/Lec02-Dataflow.pdf}
+		\end{enumerate}
+\end{enumerate}
 
-Here's a typical figure reference.  The figure is centered at the
-top of the column.  It's scaled.  It's explicitly placed.  You'll
-have to tweak the numbers to get what you want.\\
+\subsection{Related Research}
 
-% you can also use the wonderful epsfig package...
-\begin{figure}[t]
-\begin{center}
-\begin{picture}(300,150)(0,200)
-\put(-15,-30){\special{psfile = fig1.ps hscale = 50 vscale = 50}}
-\end{picture}\\
-\end{center}
-\caption{Wonderful Flowchart}
-\end{figure}
-
-This text came after the figure, so we'll casually refer to Figure 1
-as we go on our merry way.
-
-\subsection{New Subsection}
-
-It can get tricky typesetting Tcl and C code in LaTeX because they share
-a lot of mystical feelings about certain magic characters.  You
-will have to do a lot of escaping to typeset curly braces and percent
-signs, for example, like this:
-``The {\tt \%module} directive
-sets the name of the initialization function.  This is optional, but is
-recommended if building a Tcl 7.5 module.
-Everything inside the {\tt \%\{, \%\}}
-block is copied directly into the output. allowing the inclusion of
-header files and additional C code." \\
-
-Sometimes you want to really call attention to a piece of text.  You
-can center it in the column like this:
-\begin{center}
-{\tt \_1008e614\_Vector\_p}
-\end{center}
-and people will really notice it.\\
-
-\noindent
-The noindent at the start of this paragraph makes it clear that it's
-a continuation of the preceding text, not a new para in its own right.
-
-Now this is an ingenious way to get a forced space.
-{\tt Real~$*$} and {\tt double~$*$} are equivalent.
-
-Now here is another way to call attention to a line of code, but instead
-of centering it, we noindent and bold it.\\
-
-\noindent
-{\bf \tt size\_t : fread ptr size nobj stream } \\
-
-And here we have made an indented para like a definition tag (dt)
-in HTML.  You don't need a surrounding list macro pair.
-\begin{itemize}
-\item[]  {\tt fread} reads from {\tt stream} into the array {\tt ptr} at
-most {\tt nobj} objects of size {\tt size}.   {\tt fread} returns
-the number of objects read. 
-\end{itemize}
-This concludes the definitions tag.
-
-\subsection{How to Build Your Paper}
-
-You have to run {\tt latex} once to prepare your references for
-munging.  Then run {\tt bibtex} to build your bibliography metadata.
-Then run {\tt latex} twice to ensure all references have been resolved.
-If your source file is called {\tt usenixTemplate.tex} and your {\tt
-  bibtex} file is called {\tt usenixTemplate.bib}, here's what you do:
-{\tt \small
-\begin{verbatim}
-latex usenixTemplate
-bibtex usenixTemplate
-latex usenixTemplate
-latex usenixTemplate
-\end{verbatim}
-}
-
-\subsection{Last SubSection}
-
-Well, it's getting boring isn't it.  This is the last subsection
-before we wrap it up.
-
-\section{Acknowledgments}
-
-A polite author always includes acknowledgments.  Thank everyone,
-especially those who funded the work.
-
-\section{Availability}
-
-It's great when this section says that MyWonderfulApp is free software, 
-available via anonymous FTP from
-
-\begin{center}
-{\tt ftp.site.dom/pub/myname/Wonderful}\\
-\end{center}
-
-Also, it's even greater when you can write that information is also 
-available on the Wonderful homepage at
-
-\begin{center}
-{\tt http://www.site.dom/\~{}myname/SWIG}
-\end{center}
-
-Now we get serious and fill in those references.  Remember you will
-have to run latex twice on the document in order to resolve those
-cite tags you met earlier.  This is where they get resolved.
-We've preserved some real ones in addition to the template-speak.
-After the bibliography you are DONE.
-
-{\footnotesize \bibliographystyle{acm}
-\bibliography{../common/bibliography}}
-
-\theendnotes
+% TODO
 

### 2018-09-16 proposal/proposal.tex
Created another TODO task
Filled in research hypothesis with motivating examples
Updated based on facebook conversation

@@
 
-% TODO
+\begin{enumerate}
+	\item Determine data dependency tool (for C/C++) to use, conduct background research (\textit{2 weeks})
+	\item Adapt tool to identify the OpenSSL bug (\textit{6 weeks})
+	\item Test on other cases like bitcoin wallets (\textit{2 weeks})
+	\item Prepare presentation/writeup (\textit{1 week})
+	\item Integrate tool into compiler toolchain like clang (\textit{optional})
+\end{enumerate}
 
@@
 
-% TODO
+\begin{enumerate}
+	\item An automated tool can detect entropy bugs in real-world programs.
+	\item Entropy is insufficiently propagated in programs that rely on cryptography.
+\end{enumerate}
 
@@
 		\end{enumerate}
+
+	% TODO maybe have a section about cryptocurrency wallets, etc.
 \end{enumerate}
 
-\subsection{Related Research}
+\subsection{Related research}
 

### 2018-09-17 proposal/proposal.tex
Updates
updated

@@
 
-We plan to use Data Dependency tools to determine how system entropy in a given program is used by various cryptographic algorithms.
-That will allow us to identify if and when entropy is too low or is misused.
-We can release this as a tool for developers to use as part of a compiler toolchain.
-We can also use the tool on current projects that might be misusing crypto/entropy.
+\subsection{Motivation}
+One common misuse of cryptography is the misuse of entropy. Without proper random inputs, many cryptographic algorithms
+are vulnerable to basic forms of cryptanalysis. Some cryptographic schemes, such as DSA, can even disclose long-term secrets
+like the signing key when the random per-message input is low entropy, made public, or nonunique.
 
-We have two motivating examples.
-First, we would like our tool to be able to detect the issue with the Debian/OpenSSL pseudo-random number generator that was exposed in 2008.
-Second, we would like to identify potential vulnerabilities in current cryptocurrency wallet code.
+We have two motivating examples. First, we would like our tool to be able to detect the issue with the Debian/OpenSSL pseudo-random number generator that was exposed in 2008. 
+Second, we would like to identify potential vulnerabilities in current cryptocurrency wallet code as many cryptocurrency protocols rely on DSA.
+
+\subsection{Goals}
+We plan to use data dependency tools to determine how entropic inputs in a given program are used by various cryptographic algorithms. 
+That will allow us to identify if and when entropy is too low or is misused. We plan to either produce
+this as a code integration tool for developers to use as part of a compiler toolchain, or use this tool to analyze a large number of codebases
+found ``in the wild,'' such as those written by amateurs.
+
+Our tool will seek to generate a human-checkable dependency graph from source code
+using taint analysis on functions and function inputs manually specified via annotation by the developer.
+This graph will allow the developer to manually verify that entropy is used properly; a stretch goal 
+of ours would be to automate the identification of a subset of known misuses of entropy.
 
@@
 	\item Determine data dependency tool (for C/C++) to use, conduct background research (\textit{2 weeks})
-	\item Adapt tool to identify the OpenSSL bug (\textit{6 weeks})
-	\item Test on other cases like bitcoin wallets (\textit{2 weeks})
+	\item Adapt tool to identify the OpenSSL bug (\textit{5 weeks})
+	\item Test on other cases like Bitcoin wallets (\textit{3 weeks})
 	\item Prepare presentation/writeup (\textit{1 week})
@@
 
-\section{Research Hypothesis}
+\section{Research Hypotheses}
+
+We principally have two hypotheses we would like to test.
 
@@
 
+Static code analysis has a history of identifying security vulnerabilities at a source code level.
+Some examples include SQL injection, cross-site scripting exploits, and buffer overflow attacks. However,
+there has not been any attempt in the literature to statically analyze source code for cryptographic
+vulnerabilities stemming from entropy misuse, which we seek to do.
+
 % TODO maybe we need to say what exactly we found in the links instead of just dropping links?
-\subsection{Background information}
 
@@
 		\end{enumerate}
+	\item Static Program Analysis
+		\begin{enumerate}
+            \item \url{https://cs.au.dk/~amoeller/spa/spa.pdf}
+            \item \url{https://ieeexplore.ieee.org/stamp/stamp.jsp?arnumber=6859783}
+		\end{enumerate}
 
@@
 
-\subsection{Related research}
-
 % TODO

### 2018-10-29 update/update.tex
Started the project update

% TEMPLATE for Usenix papers, specifically to meet requirements of
%  USENIX '05
% originally a template for producing IEEE-format articles using LaTeX.
%   written by Matthew Ward, CS Department, Worcester Polytechnic Institute.
% adapted by David Beazley for his excellent SWIG paper in Proceedings,
%   Tcl 96
% turned into a smartass generic template by De Clarke, with thanks to
%   both the above pioneers
% use at your own risk.  Complaints to /dev/null.
% make it two column with no page numbering, default is 10 point

% Munged by Fred Douglis <douglis@research.att.com> 10/97 to separate
% the .sty file from the LaTeX source template, so that people can
% more easily include the .sty file into an existing document.  Also
% changed to more closely follow the style guidelines as represented
% by the Word sample file.

% Note that since 2010, USENIX does  not require endnotes. If you want
% foot of page notes, don't include the endnotes package in the 
% usepackage command, below.

% This version uses the latex2e styles, not the very ancient 2.09 stuff.
\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{usenix,epsfig,endnotes}

\usepackage{algpseudocode}

\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}
\begin{document}

%don't want date printed
\date{\today}

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{\Large \bf CS380S: Project Update}

%for single author (just remove % characters)
\author{
{\rm Rushi Shah}
\and
{\rm Andrew Russell}
}

\maketitle

One common misuse of cryptography is the misuse of entropy. Without proper random inputs, many cryptographic algorithms
are vulnerable to basic forms of cryptanalysis. Some cryptographic schemes, such as DSA, can even disclose long-term secrets
like the signing key when the random per-message input is low entropy, made public, or nonunique.

Given two versions $P_1, P_2$ of the same program P, we propose a relational verification for the entropy of the cryptographic values in the program. By introducing the concept of ``differential taint analysis'', we hope to produce a product program $P_1 \times P_2$ that is semantically equivalent to the sequential composition $P_1; P_2$, but such that we can prove useful safety properties of $P_1 \times P_2$ that would be difficult to prove with standard techniques on $P_1$, $P_2$, or $P_1; P_2$.

\section{Motivating Example}

Consider the following two programs $P_1$ and $P_2$ that are equivalent from a cryptographic point of view, but might have small feature changes.

\begin{algorithmic}
\Function{$P_1$}{}
\State $S \gets$ entropy
\State $\vdots$
\If {f()}
\State $c \gets$ AES(s)
\EndIf
\State $\vdots$
\EndFunction
\\
\Function{$P_2$}{}
\State $S \gets$ entropy
\State $\vdots$
\If {f()}
\State $c \gets$ AES(s)
\EndIf
\State $\vdots$
\EndFunction
\end{algorithmic}

where $f$ is some complex function that is difficult to reason about, but does not change between $P_1$ and $P_2$. Then, safety properties about the entropy of the cryptographic values in $c$ would hold in both programs or in neither program.

\section{Existing Techniques}

Static code analysis has a history of identifying security vulnerabilities at a source code level. Some examples include SQL injection, cross-site scripting exploits, and buffer overflow attacks. However, there has not been any attempt in the literature to statically analyze source code for cryptographic vulnerabilities stemming from entropy misuse, which we seek to do. Standard static code analysis techniques developed thus far are also insufficient for the analysis we would like to perform.

Under-approx tools don't exist (why do we need under approx tools?), and also difficult to reason about arbitrary functions (but they are usually irrelevant to the analysis you're performing), also can't just run taint analysis on one program because don't know what the taint results should look like (the old program gives us a correctness criteria/oracle to infer what the correct taint results would be).

\section{Our Approach}
Introduce concept of differential taint analysis here. Talk about constructing a product program.

\begin{algorithmic}
\Function{$P_1 \times P_2$}{}
\State $S_1 \gets$ entropy
\State $S_2 \gets$ entropy
\State $\vdots$
\If {f()}
\State c$_1 \gets$ AES($s_1$)
\State $c_2 \gets$ AES($s_2$)
\EndIf
\State $\vdots$
\EndFunction
\end{algorithmic}

Note that, during the construction of the product program, we need to be able to merge the if statements in the two programs. Also note that we need to demonstrate set equality between the taint for $s_1, s_2$ when they are used in the if statement.

\section{What We Hope To Show}
Talk about how we assign the output of a sufficiently entropic value to a variable in $P_1$ iff we also assign it in $P_2$. Then show that there is an equivalence in the taint propagation (prove set equality of the places where cryptographic values are used).

\section{Real World Applications}
How our analysis framework would be able to identify real world bugs. What are the benchmarks we compare against (OpenSSL, FreeBSD, etc.).

\section{Research Hypotheses}

These are the principal hypotheses we would like to test:

\begin{enumerate}
	\item An automated tool can detect entropy bugs in real-world programs.
	\item Entropy is insufficiently propagated in programs that rely on cryptography.
	\item Multiple versions of the same program can make static analysis for this domain more effective
\end{enumerate}

\section{Links}

\begin{enumerate}
	\item Debian/OpenSSL Bug 
		\begin{enumerate}
			\item \url{https://www.schneier.com/blog/archives/2008/05/random_number_b.html}
			\item \url{https://research.swtch.com/openssl}
			\item \url{https://freedom-to-tinker.com/2013/09/20/software-transparency-debian-openssl-bug/}
			\item \url{https://www.cs.umd.edu/class/fall2017/cmsc818O/papers/private-keys-public.pdf}
		\end{enumerate}
	\item Data flow
		\begin{enumerate}
			\item \url{https://en.wikipedia.org/wiki/Data-flow_analysis}
			\item \url{https://www.seas.harvard.edu/courses/cs252/2011sp/slides/Lec02-Dataflow.pdf}
		\end{enumerate}
	\item Static Program Analysis
		\begin{enumerate}
            \item \url{https://cs.au.dk/~amoeller/spa/spa.pdf}
            \item \url{https://ieeexplore.ieee.org/stamp/stamp.jsp?arnumber=6859783}
		\end{enumerate}
	\item Relational Verification:
		\begin{enumerate}
			\item \url{https://dl.acm.org/citation.cfm?id=2021319}
			\item \url{https://ac.els-cdn.com/S235222081630044X/1-s2.0-S235222081630044X-main.pdf?_tid=076a0492-9cee-4995-9710-bcb3c64b98e0&acdnat=1539815890_178849b4f14af3751e9acb03b238db4d}
			\item \url{https://www.microsoft.com/en-us/research/publication/differential-assertion-checking/}
			\item \url{https://www.microsoft.com/en-us/research/wp-content/uploads/2014/06/paper-1.pdf}
			\item \url{https://www.cs.utexas.edu/~isil/pldi16-chl.pdf}
		\end{enumerate}
	\item Projects to analyze 
		\begin{enumerate}
			\item OpenPGP
			\item BouncyCastle
			\item OpenSSL
			\item GnuPGP
			\item F\# SSL project with proof of correctness
			\item NQSBTLS
			\item Amazon's s2n (signal to noise)
		\end{enumerate}
\end{enumerate}

% We have two motivating examples. First, we would like our tool to be able to detect the issue with the Debian/OpenSSL pseudo-random number generator that was exposed in 2008. 
% Second, we would like to identify potential vulnerabilities in current cryptocurrency wallet code as many cryptocurrency protocols rely on DSA.

% \subsection{Goals}
% We plan to use data dependency tools to determine how entropic inputs in a given program are used by various cryptographic algorithms. 
% That will allow us to identify if and when entropy is too low or is misused. We plan to either produce
% this as a code integration tool for developers to use as part of a compiler toolchain, or use this tool to analyze a large number of codebases
% found ``in the wild,'' such as those written by amateurs.

% Our tool will seek to generate a human-checkable dependency graph from source code
% using taint analysis on functions and function inputs manually specified via annotation by the developer.
% This graph will allow the developer to manually verify that entropy is used properly; a stretch goal 
% of ours would be to automate the identification of a subset of known misuses of entropy.

% TODO

\end{document}

### 2018-10-29 update/update.tex
Continued working on update, left TODO for things that are blatantly missing

@@
 By introducing the concept of ``differential taint analysis'', we hope to produce a product program $P_1 \times P_2$ that is semantically equivalent to the sequential composition $P_1; P_2$, but such that we can prove useful safety properties of $P_1 \times P_2$ that would be difficult to prove with standard techniques on $P_1$, $P_2$, or $P_1; P_2$.
+The safety property we are the most interested in proving is that cryptographic values rely on sufficiently many bits of entropy when they are used.
 
-\section{Motivating Example}
+\section{Example Formalization}
 
@@
 where $f$ is some complex function that is difficult to reason about, but does not change between $P_1$ and $P_2$.
-Then, safety properties about the entropy of the cryptographic values in $c$ would hold in both programs or in neither program.
+Then, clearly, our safety properties about the entropy of the cryptographic values in $c$ would hold in both programs or in neither program.
 
-\section{Existing Techniques}
+\section{Existing Techniques, DTA Motivation}
 
@@
 
-Under-approx tools don't exist (why do we need under approx tools?), and also difficult to reason about arbitrary functions (but they are usually irrelevant to the analysis you're performing), also can't just run taint analysis on one program because don't know what the taint results should look like (the old program gives us a correctness criteria/oracle to infer what the correct taint results would be).
+Standard taint analysis will under/overapproximate the taint, so we may not be able to use it to directly prove our safety property on the target program.
+Thus, we can leverage the approach of relational verification.
+
+Relational verification aims to solve this problem by proving relational properties between two programs.
+However, the naive relational verification approach of verifying each program individually is infeasible given arbitrarily difficult-to-reason-about functions even if those functions do not change between versions of the program.
+One approach is to underapproximate the taint on one program and overapproximate the taint on another program to verify the set equality of taint sources, but this still falls prey to the approximation of taint analysis given arbitrarily complex functions.
+Also, current tooling does not widely support underapproximating taint analysis.
+
+Finally, even with a sufficiently precise taint analysis, it is unclear how to determine the ``correct'' set of sources a cryptographic value should rely on at any point in the program.
+For example, some programs will use weak entropy on system startup, and increase the entropy in those values, later on.
+Thus, by running our analysis on two versions of the program, we can use one version as an oracle for the correctness criteria of the other program.
+In other words, we can use it to infer what the set of taint results should be for any variable at any point in the program.
 
 \section{Our Approach}
-Introduce concept of differential taint analysis here. Talk about constructing a product program.
+Given the two programs in the example, we propose generating a product program similar to the following:
 
@@
 Note that, during the construction of the product program, we need to be able to merge the if statements in the two programs.
-Also note that we need to demonstrate set equality between the taint for $s_1, s_2$ when they are used in the if statement.
 
-\section{What We Hope To Show}
-Talk about how we assign the output of a sufficiently entropic value to a variable in $P_1$ iff we also assign it in $P_2$.
-Then show that there is an equivalence in the taint propagation (prove set equality of the places where cryptographic values are used).
+We would like to demonstrate set equality between the taint for $s_1, s_2$ when they are used in the if statement, which amounts to demonstrating the equivalence of the taint propagation through the program.
+We would also like to demonstrate that we assign the output of a sufficiently entropic value to a variable in $P_1$ if and only if we also assign it that sufficiently entropic value in $P_2$.
+
+Proving these two properties (set-equality and equivalence of assignments) allows us to claim that $P_1$ has the safety property if and only if $P_2$ has the safety property.
 
 \section{Real World Applications}
-How our analysis framework would be able to identify real world bugs. What are the benchmarks we compare against (OpenSSL, FreeBSD, etc.).
+
+A verification tool like this could be especially useful for a maintainer of a project that relies on cryptographic values.
+For example, if she audits her code once to confirm that it uses entropy properly, she can use differential taint analysis on future commits to confirm that the new code does not introduce this class of bugs.
+
+How our analysis framework would be able to identify real world bugs. What are the benchmarks we compare against (OpenSSL, FreeBSD, etc.). What are the projects we will evaluate the tool on. %TODO
 

### 2018-10-30 update/update.tex
Clarified existing techniques

@@
 
-\section{Example Formalization}
+\section{Example}
 
@@
 \Function{$P_1$}{}
-\State $S \gets$ entropy
+\State $s \gets$ entropy
 \State $\vdots$
-\If {f()}
+\If {$p$}
 \State $c \gets$ AES(s)
@@
 \Function{$P_2$}{}
-\State $S \gets$ entropy
+\State $s \gets$ entropy
 \State $\vdots$
-\If {f()}
+\If {$p$}
 \State $c \gets$ AES(s)
@@
 
-where $f$ is some complex function that is difficult to reason about, but does not change between $P_1$ and $P_2$.
-Then, clearly, our safety properties about the entropy of the cryptographic values in $c$ would hold in both programs or in neither program.
+where $p$ is some complex predicate that is difficult to reason about, but does not change between $P_1$ and $P_2$ (and no values that $p$ depends on change, either).
+Then, a safety property about the entropy of the cryptographic values in $c$ would hold in both programs or in neither program.
 
@@
 However, there has not been any attempt in the literature to statically analyze source code for cryptographic vulnerabilities stemming from entropy misuse, which we seek to do.
-Standard static code analysis techniques developed thus far are also insufficient for the analysis we would like to perform.
+We claim that standard static code analysis techniques developed thus far are insufficient for the analysis we would like to perform.
 
-Standard taint analysis will under/overapproximate the taint, so we may not be able to use it to directly prove our safety property on the target program.
-Thus, we can leverage the approach of relational verification.
+The sources of entropy can be tainted standard taint analysis techniques can be used to approximate how the taint propagates.
+But this approximation may be too coarse to provide useful information to the user.
+Even with a sufficiently precise taint analysis, it is unclear how to determine the ``correct'' set of sources a cryptographic value should rely on at any point in the program.
 
-Relational verification aims to solve this problem by proving relational properties between two programs.
+Verification modulo versions can be leveraged, the static analyzer is treated as opaque in an effort to cut down the number of alarms by inferring assumptions made by the original program (assumed to be correct).
+This is especially appealing in the security software development cycle because there are documented cases of entropy bugs being introduced in new versions of programs (for example, OpenSSL and FreeBSD).
+This would involve under-approximating the taint of one version of the program and over-approximating the taint of another version of the program.
+Current tooling does not widely support under-approximating taint analysis.
+
+Another existing technique is relational verification, in which relational properties between two programs are used to prove $k$-safety properties.
+In our case, we would want to show the equivalence of entropy assignment and the equivalence of the entropy propagation.
+Relational verification would proceed with the construction of a product program between versions of the program that captures the semantics of executing them sequentially, but that lends itself to standard verification techniques.
+
+% How is vmv related to relational verification?
+
+% Standard taint analysis will under/overapproximate the taint, so we may not be able to use it to directly prove our safety property on the target program.
+Thus, we can leverage the approach verification modulo versions to prove relational verification.
+
+% Relational verification aims to solve this problem by proving relational properties between two programs.
 However, the naive relational verification approach of verifying each program individually is infeasible given arbitrarily difficult-to-reason-about functions even if those functions do not change between versions of the program.
@@
 
-Finally, even with a sufficiently precise taint analysis, it is unclear how to determine the ``correct'' set of sources a cryptographic value should rely on at any point in the program.
+% Finally, even with a sufficiently precise taint analysis, it is unclear how to determine the ``correct'' set of sources a cryptographic value should rely on at any point in the program.
 For example, some programs will use weak entropy on system startup, and increase the entropy in those values, later on.
@@
 For example, if she audits her code once to confirm that it uses entropy properly, she can use differential taint analysis on future commits to confirm that the new code does not introduce this class of bugs.
+% VMV?
 
@@
 
-% TODO
-
 \end{document}

### 2018-10-30 update/update.tex
proposal updates

@@
 
+\section{Introduction}
+
 One common misuse of cryptography is the misuse of entropy. Without proper random inputs, many cryptographic algorithms
@@
 
-Given two versions $P_1, P_2$ of the same program P, we propose a relational verification for the entropy of the cryptographic values in the program.
-By introducing the concept of ``differential taint analysis'', we hope to produce a product program $P_1 \times P_2$ that is semantically equivalent to the sequential composition $P_1; P_2$, but such that we can prove useful safety properties of $P_1 \times P_2$ that would be difficult to prove with standard techniques on $P_1$, $P_2$, or $P_1; P_2$.
-The safety property we are the most interested in proving is that cryptographic values rely on sufficiently many bits of entropy when they are used.
+We plan to use data dependency tools to determine how entropic inputs in a given program are used by various cryptographic algorithms.
+That will allow us to identify if and when entropy is too low or is misused.
+We plan to either produce this as a code integration tool for developers to use as part of a compiler toolchain, or use this tool to analyze a large number of codebases found ``in the wild,'' such as those written by amateurs.
+Our tool will seek to generate a human-checkable dependency graph from source code using notions of taint analysis.
+
+For this project we consider codebases that maintain version history, in the hopes that we can employ static analysis techniques across program versions to infer more information about how entropy is being used;
+our primary goal is to identify bugs that are introduced to existing codebases.
+This may prove to be especially helpful since naive taint analysis can provide mere overapproximations of data dependency for a given program and may generate many false alarms for programmers in a professional development setting.
+
+\section{Considered Approaches}
+
+One technique that we have considered is the notion of a ``product program.'' In the most general sense, the product $P_1 \times P_2$ of two programs $P_1$ and $P_2$ is used to verify
+relations between the programs, such as equivalence. Product programs have also been used to analyze different runs of the same program. 
+The product program $P_1 \times P_2$ is semantically equivalent to the sequential composition $P_1; P_2$, but such that we can prove useful safety properties of $P_1 \times P_2$ that would be difficult to prove with standard techniques on $P_1$, $P_2$, or $P_1; P_2$.
+
+In our setting, we are not so much interested in proving safety properties of a program $P$ but instead wish to prove that cryptographic values rely on sufficiently many bits of entropy when they are used via taint analysis.
+The notion of product programs helps us introduces the concept of ``differential taint analysis'' which will give u
+
+Given two versions $P_1, P_2$ of the same program P, we propose a relational verification for the entropy of the cryptographic values in the program. By  
+we hope to produce a The safety property we are the most interested in proving is that
 

### 2018-12-10 CPAchecker/bench1/sequential-basic.c
Preserving/nonpreserving test

@@ -5,8 +5,6 @@
 
 #include "rng.h"
 
-#include "assert.h"
-
 #define KEY_SIZE 128
 
 int main(int argc, char** argv) {
@@ -25,7 +23,7 @@ int main(int argc, char** argv) {
         printf("Failed to open /dev/urandom with errno %d\n", errno);
         return -1;
     } else {
-        ssize_t result = read(urandomFd, seed, key_SIZE);
+        ssize_t result = read(urandomFd, seed, KEY_SIZE);
         if (result != KEY_SIZE) {
             printf("Failed to read %d bytes from /dev/urandom as requested, with errno %d\n", KEY_SIZE, errno);
             return -1;
@@ -45,8 +43,9 @@ int main(int argc, char** argv) {
         }
         close(urandomFd2);
         // comment this line out for bad things!
-        // rng2.update_seed(&rng2, seed2, KEY_SIZE);
+        rng2.update_seed(&rng2, seed2, KEY_SIZE);
     }
+    assert(rng.seed == rng2.seed); // this assertion is what I don't really understand
 
     pid_t pid = getpid();
     pid_t pid2 = getpid();
@@ -55,21 +54,20 @@ int main(int argc, char** argv) {
     for (int i = 0; i < 4; i++) {
         pid_seed[i] = (pid >> (24 - 8*i)) & 0xFF;
     }
-    for (int i2 = 0; i2 < 4; i2++) {
-        pid_seed2[i] = (pid2 >> (24 - 8*i2)) & 0xFF;
+    for (int i = 0; i < 4; i++) {
+      pid_seed2[i] = (pid2 >> (24 - 8*i)) & 0xFF;
     }
     rng.update_seed(&rng, pid_seed, 4);
     rng2.update_seed(&rng2, pid_seed2, 4);
-
+    
     rng.gen_random(&rng, rand_bytes, sizeof rand_bytes);
-    rng.gen_random(&rng2, rand_bytes2, sizeof rand_bytes2);
+    rng2.gen_random(&rng2, rand_bytes2, sizeof rand_bytes2);
     for(int i = 0; i < sizeof rand_bytes; i++) {
         printf("%.2x", rand_bytes[i]);
     }
-    for(int i2 = 0; i2 < sizeof rand_bytes2; i2++) {
+    for(int i = 0; i < sizeof rand_bytes2; i++) {
         printf("%.2x", rand_bytes2[i]);
     }
     printf("\n");
-
-    return 0;
+    printf("\n");
 }

### 2018-12-10 CPAchecker/output/Counterexample.1.c

@@ -1,30 +1,0 @@
-void __assert_fail(const char *__assertion, const char *__file, unsigned int __line, const char *__function);
-void __assert_perror_fail(int __errnum, const char *__file, unsigned int __line, const char *__function);
-void __assert(const char *__assertion, const char *__file, int __line);
-int p1();
-int p2();
-int A1 = 101;
-int main();
-extern void __VERIFIER_error(void);
-int main_0();
-int main_0() {
-  int rng1;
-  int rng2;
-  int __CPAchecker_TMP_0;
-  __CPAchecker_TMP_0 = p1();
-  __CPROVER_assume(__CPAchecker_TMP_0 == 0);
-  int seed1 = A1;
-  int seed2 = A1;
-  void __CPAchecker_TMP_1;
-  __CPROVER_assume(seed1 == seed2);
-  0;
-  0;
-  int __CPAchecker_TMP_2;
-  __CPAchecker_TMP_2 = p2();
-  __CPROVER_assume(__CPAchecker_TMP_2 == 0);
-  rng1 = seed1;
-  void __CPAchecker_TMP_3;
-  __CPROVER_assume(!(rng1 == rng2));
-  __VERIFIER_error(); // target state
-  __assert_fail("rng1 == rng2", "bench1/simplified-product-incorrect.c", 28, "__PRETTY_FUNCTION__");
-}

### 2018-12-10 CPAchecker/preserving.c

@@ -1,0 +1,14 @@
+#include "assert.h"
+
+extern int preserving(int x, int y);
+extern int non_preserving(int x, int y);
+
+int main() {
+  assert(
+	 preserving(preserving(1, 2), 3)
+	 ==
+	 preserving(non_preserving(1, 2), 3)
+	 );
+  return 0;
+}
+

### 2018-12-12 paper/paper.tex
presentation and paper start

--- 2018-10-30 update/update.tex
@@
 %make title bold and 14 pt font (Latex default is non-bold, 16 pt)
-\title{\Large \bf CS380S: Project Update}
+\title{\Large \bf New Static Analysis Techniques to Detect Entropy Failure Vulnerabilities}
 
@@
 are vulnerable to basic forms of cryptanalysis. Some cryptographic schemes, such as DSA, can even disclose long-term secrets
-like the signing key when the random per-message input is low entropy, made public, or nonunique.
+such as the signing key when the random per-message input is low entropy, made public, or nonunique.
 
@@
 That will allow us to identify if and when entropy is too low or is misused.
-We plan to either produce this as a code integration tool for developers to use as part of a compiler toolchain, or use this tool to analyze a large number of codebases found ``in the wild,'' such as those written by amateurs.
-Our tool will seek to generate a human-checkable dependency graph from source code using notions of taint analysis.
+We plan to produce this as a code integration tool for developers to use as part of a compiler toolchain.
+Our tool will seek to generate an error report from source code using notions of taint analysis.
 
@@
 
+\section{Preliminaries}
+
+\subsection{Taint Analysis}
+
+\subsection{Predicate Abstraction}
+
+\subsection{$k$-safety properties}
+
+\subsection{Product Programs}
+
+\section{Our Approach}
+
+\subsection{Instrumenting Programs for Predicate Abstraction}
+
+\subsection{Naive Product Program Inference Rules}
+
+\subsection{Taint-augmented Inference Rules}
+
+\subsection{Putting it all together}
+
+\section{Results}
+
+\section{Conclusion}
+
+\section{Future Work}
+
 \section{Considered Approaches}
 
-One technique that we have considered is the notion of a ``product program.'' In the most general sense, the product $P_1 \times P_2$ of two programs $P_1$ and $P_2$ is used to verify
+\subsection{Taint Analysis}
+Static code analysis has a history of identifying security vulnerabilities at a source code level.
+Some examples include SQL injection, cross-site scripting exploits, and buffer overflow attacks.
+However, there has not been any attempt in the literature to statically analyze source code for cryptographic vulnerabilities stemming from entropy misuse, which we seek to do.
+We claim that standard static code analysis techniques developed thus far are insufficient for the analysis we would like to perform.
+The sources of entropy can be tainted standard taint analysis techniques can be used to approximate how the taint propagates.
+But this approximation may be too coarse to provide useful information to the user.
+Even with a sufficiently precise taint analysis, it is unclear how to determine the ``correct'' set of sources a cryptographic value should rely on at any point in the program.
+
+\subsection{Differential Taint Analysis}
+
+Since we want a stronger notion of taint analysis than a naive approach, we intend to introduce and formalize the concept of ``differential taint analysis." Differential
+taint analysis, in the ideal, will give us more information than simple taint analysis by leveraging 
+the original version of the program as a source of ground truth. This additional information could include reducing programmer annotation work or reduce false positives between versions
+(since taint analysis overapproximates data flow). By viewing the original version of the program as bug-free, we reduce the number of code audits required:
+only the original version of the code would need to be audited, and future versions could rely on our tool to automatically detect bugs that were introduced by refactors.
+
+\subsection{Product Programs}
+
+One technique that we have considered toward the construction of differential taint analysis is the notion of a ``product program.'' In the most general sense, the product $P_1 \times P_2$ of two programs $P_1$ and $P_2$ is used to verify
 relations between the programs, such as equivalence. Product programs have also been used to analyze different runs of the same program. 
@@
 
-In our setting, we are not so much interested in proving safety properties of a program $P$ but instead wish to prove that cryptographic values rely on sufficiently many bits of entropy when they are used via taint analysis.
-The notion of product programs helps us introduces the concept of ``differential taint analysis'' which will give u
+In our setting, we are not so much interested in proving safety properties of a program $P$ but instead wish to prove that cryptographic values rely on sufficiently many bits of entropy when they are used via taint analysis,
+which is why we cannot simply leverage existing off-the-shelf product program analysis techniques.
 
-Given two versions $P_1, P_2$ of the same program P, we propose a relational verification for the entropy of the cryptographic values in the program. By  
-we hope to produce a The safety property we are the most interested in proving is that
-
-\section{Example}
+\subsubsection{Example}
 
@@
 
-\section{Existing Techniques, DTA Motivation}
-
-Static code analysis has a history of identifying security vulnerabilities at a source code level.
-Some examples include SQL injection, cross-site scripting exploits, and buffer overflow attacks.
-However, there has not been any attempt in the literature to statically analyze source code for cryptographic vulnerabilities stemming from entropy misuse, which we seek to do.
-We claim that standard static code analysis techniques developed thus far are insufficient for the analysis we would like to perform.
-
-The sources of entropy can be tainted standard taint analysis techniques can be used to approximate how the taint propagates.
-But this approximation may be too coarse to provide useful information to the user.
-Even with a sufficiently precise taint analysis, it is unclear how to determine the ``correct'' set of sources a cryptographic value should rely on at any point in the program.
-
-Verification modulo versions can be leveraged, the static analyzer is treated as opaque in an effort to cut down the number of alarms by inferring assumptions made by the original program (assumed to be correct).
+\subsection{Verification Modulo Versions}
+Verification modulo versions is an idea quite similar to product programs. VMV attempts to identify abstract regressions and relative correctness between two versions of a program.
+In VMV, the static analyzer is treated as opaque in an effort to cut down the number of alarms by inferring assumptions made by the original program (assumed to be correct).
 This is especially appealing in the security software development cycle because there are documented cases of entropy bugs being introduced in new versions of programs (for example, OpenSSL and FreeBSD).
@@
 
+\subsection{Relational Verification}
 Another existing technique is relational verification, in which relational properties between two programs are used to prove $k$-safety properties.
+A $k$-safety property is a safety property of a program that requires
+reasoning about the relationships between $k$ different runs of a program (or $k$ programs). For example, transitivity is a $k$-safety property.
 In our case, we would want to show the equivalence of entropy assignment and the equivalence of the entropy propagation.
-Relational verification would proceed with the construction of a product program between versions of the program that captures the semantics of executing them sequentially, but that lends itself to standard verification techniques.
+Relational verification utilitzes product programs; RV constructs a product program between versions of the program that captures the semantics of executing them sequentially, but that lends itself to standard verification techniques.
+
+\subsection{Differential Assertion Checking}
+
+The last approach that we have considered is differential assertion checking. Differential assertion checking seeks to (statically) prove the relative correctness between two similar programs,
+with a significantly lower cost than ensuring absolute correctness. That is, DAC allows is to answer if there are conditions under which a program $P_1$ passes an assertion check but
+a program $P_2$ fails. The techniques utilized current DAC work may be of interest, but it seems they rely on an a priori mapping between the two programs, which is something that
+relational verification work seeks to answer. Thus DAC will be considered further after the idea of product programs is exhausted.
 
@@
 
-How our analysis framework would be able to identify real world bugs. What are the benchmarks we compare against (OpenSSL, FreeBSD, etc.). What are the projects we will evaluate the tool on. %TODO
+Our tool should be able to detect the infamous Debian OpenSSL entropy bug.
+We will also test our tool on other libraries that rely on entropy, such as Amazon's signal2noise, LibreSSL, GnuTLS, and GnuPG,
+which are not known to contain entropy bugs (and likely do not), but we can test on versions in which we introduce our own bugs.
 
@@
 	\item An automated tool can detect entropy bugs in real-world programs.
-	\item Entropy is insufficiently propagated in programs that rely on cryptography.
-	\item Multiple versions of the same program can make static analysis for this domain more effective
+	\item Entropy is insufficiently propagated in programs that rely on cryptography, or entropy propagation follows nontrivial code paths (due to error handling or other control flow).
+	\item Multiple versions of the same program can make static analysis for this domain more effective by lowering costs (computational and programmer) or providing more fine-grained information.
 \end{enumerate}
+
+\section{Future Work}
+
+As bad data flow can be a source of security bugs other than simply (lack of) entropy propagation (for example, 
+address disclosure and use-after-free bugs in browsers), we believe that our tool could also be applied to these settings to help prevent
+programmers from introducing security vulnerabilities into their projects.
 

### 2018-12-12 inference_rules/inference_rules.tex
Inference rules

\documentclass{article}
\usepackage{proof}

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{stmaryrd}

\newcommand{\cross}{\otimes{}}

\begin{document}

We are using a state-of-the-art off-the-shelf predicate abstraction tool called CPAChecker. Sequentially composing $V_1$ and $V_2$ as $V_1\ ;\ V_2$ is a valid input to CPAChecker, but it might have trouble verifying the assertions. Constructing a product program by synchronizing the statements of $V_1$ and $V_2$ gives the predicate abstraction tool more power to reason about the composition. This synchronized approach is semantically equivalent to the sequential composition of the two programs. Although it is easier for CPAChecker to prove the assertions, the synchronized approach can also lead to an exponential blowup in the product program size in the presence of control flow branching or loops. Thus we outline three possible approaches:

\begin{enumerate}
	\item Sequential composition 
	\item Synchronized composition
	\item Heuristic-optimized synchronized composition
\end{enumerate}

We will work with a simple set of standard semantics:

\[
	\begin{array}{l l c l}
		\emph{Statement} & S & := & 
			\ \ A \\
			& & & |~ S_1\ ;\ S_2 \\
			& & & |~ \emph{if }p\emph{ then } S_1 \emph{ else } S_2 \\
			& & & |~ \emph{while }p\emph{ } S \\
		\emph{Predicate} & p &:=& \top ~~|~~ \bot ~~|~~ A ~~|~~ \lnot p ~~|~~ p \odot p \\
		\emph{Operator} & \odot &:=& \land ~~|~~ \lor \\
	\end{array}
\]

\subsection{Sequential Composition}

	Consider the following code snippet and :

	\begin{algorithm}[H]
		\begin{algorithmic}
			\If {complicatedFunction()}
			    \State $x \gets legitRandomness()$
			\Else
			    \State $x \gets getPID()$
			\EndIf
		\end{algorithmic}
	\end{algorithm}

	and it's self-composiiton

	\begin{algorithm}[H]
		\begin{algorithmic}
			\If {complicatedFunction()}
			    \State $x \gets legitRandomness()$
			\Else
			    \State $x \gets getPID()$
			\EndIf
			\If {complicatedFunction()}
			    \State $x' \gets legitRandomness()$
			\Else
			    \State $x' \gets getPID()$
			\EndIf
			\State $assert(x == x')$
		\end{algorithmic}
	\end{algorithm}

	\begin{algorithm}[H]
		\begin{algorithmic}
			\If {complicatedFunction()}
			    \State $x \gets legitRandomness()$
			    \If {complicatedFunction()}
    			    \State $x' \gets legitRandomness()$
    			\Else
    			    \State $x' \gets getPID()$
    			\EndIf
    			\State $assert(x == x')$
			\Else
			    \State $x \gets getPID()$
			    \If {complicatedFunction()}
				    \State $x' \gets legitRandomness()$
				\Else
				    \State $x' \gets getPID()$
				\EndIf
				\State $assert(x == x')$
			\EndIf
		\end{algorithmic}
	\end{algorithm}

\subsection{Synchronized Composiiton}

	Here are our naive inference rules. These represent our approach for synchronized composition

	\[
		\begin{array}{c}
			\infer[]{A_1 \cross A_2 \leadsto A_1\ ;\ A_2}{} \\ \\ \\
			\infer[]{S_1 \cross S_2 \leadsto P}{
				S_2 \cross S_1 \leadsto P
			} \\ \\ \\
			\infer[]{if(p)\ then\ S_1\ else\ S_2 \cross S \leadsto P}{
				\text{$S_1 \cross S \leadsto S_1'$} \\
				\text{$S_2 \cross S \leadsto S_2'$} \\
				\text{$P = if(p)\ then\ S_1'\ else\ S_2'$}
			} \\ \\ \\
			\infer[]{while(p_1)\ S_1 \cross while(p_2)\ S_2 \leadsto P_0\ ;\ P_1\ ;\ P_2}{
				\text{$P_0 = while(p_1 \land p_2)\ S_1\ ;\ S_2$} \\
				\text{$P_1 = while(p_1)\ S_1$} \\
				\text{$P_2 = while(p_2)\ S_2$}
			}
		\end{array}
	\]

\subsection{Heuristic-optimized synchronized composition}

	Now we can add an environment $\Gamma$ such that $\Gamma \vdash S$ if there is some atomic statement $A$ in $S$ such that $A$ references a tainted variable. Conversely $\Gamma \nvdash S$ if for every variable $v$ in $S$, $v$ is not tainted by a source of genuine randomness.

	Our key insight is that the following inference rule can be used to optimize the synchronization

	\[
		\begin{array}{c}
			\infer[]{\Gamma \vdash S_1 \cross S_2 \leadsto S_1\ ;\ S_2}{
				\Gamma \nvdash S_1 \\
				\Gamma \nvdash S_2
			} \\
		\end{array}
	\]

	Our previous inference rules can also be trivially modified with $\Gamma$. This would give us the following set of inference rules:

	\[
		\begin{array}{c}
			\infer[]{\Gamma \vdash S_1 \cross S_2 \leadsto S_1\ ;\ S_2}{
				\Gamma \nvdash S_1 \\
				\Gamma \nvdash S_2
			} \\ \\ \\
			\infer[]{\Gamma \vdash A_1 \cross A_2 \leadsto A_1\ ;\ A_2}{} \\ \\ \\
			\infer[]{\Gamma \vdash S_1 \cross S_2 \leadsto P}{
				\Gamma \vdash S_2 \cross S_1 \leadsto P
			} \\ \\ \\
			\infer[]{\Gamma \vdash if(p)\ then\ S_1\ else\ S_2 \cross S \leadsto P}{
				\text{$\Gamma \vdash S_1 \cross S \leadsto S_1'$} \\
				\text{$\Gamma \vdash S_2 \cross S \leadsto S_2'$} \\
				\text{$P = if(p)\ then\ S_1'\ else\ S_2'$}
			} \\ \\ \\
			\infer[]{\Gamma \vdash while(p_1)\ S_1 \cross while(p_2)\ S_2 \leadsto P_0\ ;\ P_1\ ;\ P_2}{
				\text{$P_0 = while(p_1 \land p_2)\ S_1\ ;\ S_2$} \\
				\text{$P_1 = while(p_1)\ S_1$} \\
				\text{$P_2 = while(p_2)\ S_2$}
			}
		\end{array}
	\]

	This is only a heuristic because it is not clear that the heuristic-optimized synchronized composition invariably gives CPAChecker as much power as the synchronized version. For example, consider the following counterexample and its self-composition:

	\begin{algorithm}[H]
		\begin{algorithmic}
			\If {flag}
			    \State $x \gets complicatedFunction1()$
			\Else
			    \State $x \gets complicatedFunction2()$
			\EndIf
			\If {$x == 7$}
				\State $y \gets taint()$
			\EndIf
			\State $sink(y)$
		\end{algorithmic}
	\end{algorithm}

	The theorem prover with our optimized rules would sequentially compose the initial if-statements because none of their program variables in the initial if statement are tainted. This would make it difficult or impossible for CPAChecker to prove the assertion (it would involve reasoning about the complicated functions). In contrast, the synchronized version might be able to relationally verify that if $y$ gets taint in $V_1$ then $y$ will get taint in $V_2$ while still treating the complicated functions as uninterpreted (it would still reason about the flag in version one and the flag in version two). % Not sure about my claims about the complicated functions and uninterpreted reasoning, etc. in this paragraph.

\end{document}

### 2018-12-12 paper/paper.tex
bib
paper branch

@@
 
+In this section we describe the background necessary to understand our new approach.
+
+Test reference: \cite{GKW18}
+
 \subsection{Taint Analysis}
@@
 \section{Our Approach}
+
+\subsection{Naive Approaches}
+
+Before we cover our approach to this problem we will briefly discuss a number of approaches that do not work, but build to our final result.
+
+A first attempt to solve this problem might include running taint analysis on each version $P_1$ and $P_2$ of the program. However, since
+taint analysis merely provides an overapproximation to this problem, this approach is necessarily unsound. Instead, we would like to view
+this as a $2$-safety property:
 
@@
 
+\bibliographystyle{alpha}
+\bibliography{refs}
+
 % We have two motivating examples. First, we would like our tool to be able to detect the issue with the Debian/OpenSSL pseudo-random number generator that was exposed in 2008. 

### 2018-12-12 CPAchecker/difficult.c
CPAChecker experiments

@@ -1,0 +1,31 @@
+#include "assert.h"
+
+extern int f();
+extern int g();
+
+int main() {
+  int x1;
+  int x2;
+
+  if(f()) {
+    x1 = 1;
+    if(g()) {
+      x2 = 1;
+    }
+    else {
+      x2 = 2;
+    }
+  }
+  else {
+    x1 = 2;
+    if(g()) {
+      x2 = 1;
+    }
+    else {
+      x2 = 2;
+    }
+  }
+    
+  assert(x1 == x2);
+  return (0);
+}

### 2018-12-12 CPAchecker/heuristic-synchronized.c

@@ -1,0 +1,45 @@
+#include "assert.h"
+
+extern int f();
+extern int g1();
+extern int g2();
+extern int taint();
+
+int main() {
+  int x1;
+  int x2;
+  int y1;
+  int y2;
+  
+  if(f()) {
+    x1 = g1();
+    if(f()) {
+      x2 = g1();
+    }
+    else {
+      x2 = g2();
+    }
+    if(x2 == 7) {
+      y2 = taint();
+    }
+  }
+  else {
+    x1 = g2();
+    if(f()) {
+      x2 = g1();
+    }
+    else {
+      x2 = g2();
+    }
+    if(x2 == 7) {
+      y2 = taint();
+    }
+  }
+  
+  if(x1 == 7) {
+    y1 = taint();
+  }
+  
+  assert(y1 == y2);
+  return (0);
+}

### 2018-12-12 CPAchecker/heuristic.c

@@ -1,0 +1,41 @@
+#include "assert.h"
+
+extern int f();
+extern int g1();
+extern int g2();
+extern int taint();
+extern int getPID();
+
+int main() {
+  int x1;
+  int x2;
+  int y1;
+  int y2;
+  
+  if(f()) {
+    x1 = g1();
+  }
+  else {
+    x1 = g2();
+  }
+  if(x1 == 7) {
+    y1 = taint();
+  } else {
+    y1 = getPID();
+  }
+
+  if(f()) {
+    x2 = g1();
+  }
+  else {
+    x2 = g2();
+  }
+  if(x2 == 7) {
+    y2 = taint();
+  } else {
+    y2 = getPID();
+  }
+  
+  assert(y1 == y2);
+  return (0);
+}

### 2018-12-13 paper/paper.tex
paper work

--- 2018-12-12 paper/paper.tex
@@
 
+\usepackage{proof}
+
+\usepackage{amssymb}
+\usepackage{amsmath}
+\usepackage{amsthm}
+\usepackage{mathtools}
+\usepackage{algorithm}
 \usepackage{algpseudocode}
+\usepackage{stmaryrd}
+
+\newcommand{\cross}{\otimes{}}
 
@@
 
+\subsection{Toy Language}
+
+We will build our tool for a language consisting of a small set of semantic rules (Figure \ref{fig:semrules}).
+
+\begin{figure}
+\label{fig:semrules}
+\caption{}
+\[
+	\begin{array}{l l c l}
+		\emph{Statement} & S & := & 
+			\ \ A \\
+			& & & |~ S_1\ ;\ S_2 \\
+			& & & |~ \emph{if }p\emph{ then } S_1 \emph{ else } S_2 \\
+			& & & |~ \emph{while }p\emph{ } S \\
+		\emph{Predicate} & p &:=& \top ~~|~~ \bot ~~|~~ A ~~|~~ \lnot p ~~|~~ p \odot p \\
+		\emph{Operator} & \odot &:=& \land ~~|~~ \lor \\
+	\end{array}
+\]
+\end{figure}
+
+Our results can be extended in a straightforward manner to an industrial standard language such as C.
+
 \subsection{Taint Analysis}
+
+Taint analysis is a standard program analysis technique, typically used to detect security vulnerabilities in either a static or dynamic manner. 
+We restrict our focus to the static variant. Taint analysis takes as input an initial assignment of some taint sets $\{\tau_i\}$ to 
+a sources (program variables) $\{\sigma_i\}$, as well as a list of sinks $\{\zeta_i\}$ (function definitions). 
+Taint is propagated through program statements using the semantics of the programming language;
+for example, a variable $z = x + y$ will be assigned taint set $\tau_x \cup \tau_y$. The taint analysis outputs the taint set for each invocation
+of the sinks $\{\zeta_i\}$.
+
+An example of static taint analysis is to determine if unsanitized user inputs are ever provided to vulnerable functions, such as SQL commands or
+a webpage templating function. In our case we wish to ensure that entropy-sensitive inputs to cryptographic functions (sinks) can be
+traced back to a high-entropy source such as *nix's \texttt{/dev/urandom}.
+
+\subsection{$k$-safety properties}
+
+As opposed to a safety property of a program (for example, ``variable $x$ is always positive'' or ``pointer $p$ is never null''), which requires
+the absence of errors in a single program trace, a $k$-safety property of a program requires the absence of erroneous interactions between $k$
+traces of the same program \cite{sousa2016cartesian}. A program property such as symmetry is a $2$-safety property.
 
@@
 
-\subsection{$k$-safety properties}
+Predicate abstraction is a specialized form of abstract interpretation that can be used for checking $1$-safety properties of programs
+\cite{flanagan2002predicate}.
 
@@
 
+Want to use differential view since we can get more out of this. Unsound taint analysis example.
+
+\subsection{Differential Taint Analysis}
+
+Description of differential taint analysis.
+
 \subsection{Instrumenting Programs for Predicate Abstraction}
+
+Description of instrumentation algorithm.
 
@@
 
+Description of product program inference rules approach.
+
+\begin{figure}
+    \label{fig:infrules1}
+    \caption{Basic Inference Rules}
+    \[
+		\begin{array}{c}
+			\infer[]{A_1 \cross A_2 \leadsto A_1\ ;\ A_2}{} \\ \\ \\
+			\infer[]{S_1 \cross S_2 \leadsto P}{
+				S_2 \cross S_1 \leadsto P
+			}
+        \end{array}
+    \]
+\end{figure}
+
+\begin{figure}
+    \label{fig:infrules2}
+    \caption{Conditional and Loop Inference Rules}
+    \[
+		\begin{array}{c}
+			\infer[]{if(p)\ then\ S_1\ else\ S_2 \cross S \leadsto P}{
+				\text{$S_1 \cross S \leadsto S_1'$} \\
+				\text{$S_2 \cross S \leadsto S_2'$} \\
+				\text{$P = if(p)\ then\ S_1'\ else\ S_2'$}
+			} \\ \\ \\
+			\infer[]{while(p_1)\ S_1 \cross while(p_2)\ S_2 \leadsto P_0\ ;\ P_1\ ;\ P_2}{
+				\text{$P_0 = while(p_1 \land p_2)\ S_1\ ;\ S_2$} \\
+				\text{$P_1 = while(p_1)\ S_1$} \\
+				\text{$P_2 = while(p_2)\ S_2$}
+			}
+        \end{array}
+    \]
+\end{figure}
+
 \subsection{Taint-augmented Inference Rules}
+
+Why the naive approach doesn't work (intractability). Introduce heuristic-based approach.
+
+\begin{figure}
+    \label{fig:auginfrules1}
+    \caption{Augmented Inference Rules}
+    \[
+		\begin{array}{c}
+			\infer[]{\Gamma \vdash S_1 \cross S_2 \leadsto S_1\ ;\ S_2}{
+				\Gamma \nvdash S_1 \\
+				\Gamma \nvdash S_2
+			} \\
+		\end{array}
+    \]
+\end{figure}
+
+\begin{figure}
+    \label{fig:auginfrules2}
+    \caption{Augmented Inference Rules}
+    \[
+		\begin{array}{c}
+			\infer[]{\Gamma \vdash A_1 \cross A_2 \leadsto A_1\ ;\ A_2}{} \\ \\ \\
+			\infer[]{\Gamma \vdash S_1 \cross S_2 \leadsto P}{
+				\Gamma \vdash S_2 \cross S_1 \leadsto P
+			} \\ \\ \\
+			\infer[]{\Gamma \vdash if(p)\ then\ S_1\ else\ S_2 \cross S \leadsto P}{
+				\text{$\Gamma \vdash S_1 \cross S \leadsto S_1'$} \\
+				\text{$\Gamma \vdash S_2 \cross S \leadsto S_2'$} \\
+				\text{$P = if(p)\ then\ S_1'\ else\ S_2'$}
+			} \\ \\ \\
+			\infer[]{\Gamma \vdash while(p_1)\ S_1 \cross while(p_2)\ S_2 \leadsto P_0\ ;\ P_1\ ;\ P_2}{
+				\text{$P_0 = while(p_1 \land p_2)\ S_1\ ;\ S_2$} \\
+				\text{$P_1 = while(p_1)\ S_1$} \\
+				\text{$P_2 = while(p_2)\ S_2$}
+			}
+		\end{array}
+    \]
+\end{figure}
 
@@
 
-\section{Results}
+Description of algorithm using the two subroutines above.
+
+\section{Evaluation}
 
@@
 
+In this work we present a new static analysis technique to aid in the prevention of entropy-misuse security vulnerabilities, such as those
+found in the Debian OpenSSL and FreeBSD projects.
+
 \section{Future Work}
+
+Verify the following conjectures. Prove that a heuristic-based approach is no less powerful than a full product program approach.
+Run this tool on projects in the wild to determine how widespread entropy-misuse bugs are among software projects ``in the wild.''
 

### 2018-12-13 paper/paper.tex
updates
taint analysis

@@
 
+\begin{abstract}
+ Test abstract
+\end{abstract}
+
 \section{Introduction}
@@
 Taint analysis is a standard program analysis technique, typically used to detect security vulnerabilities in either a static or dynamic manner. 
-We restrict our focus to the static variant. Taint analysis takes as input an initial assignment of some taint sets $\{\tau_i\}$ to 
-a sources (program variables) $\{\sigma_i\}$, as well as a list of sinks $\{\zeta_i\}$ (function definitions). 
-Taint is propagated through program statements using the semantics of the programming language;
-for example, a variable $z = x + y$ will be assigned taint set $\tau_x \cup \tau_y$. The taint analysis outputs the taint set for each invocation
-of the sinks $\{\zeta_i\}$.
+We restrict our focus to the static variant. We define a taint analysis algorithm as follows:
+
+\begin{itemize}
+    \item $\Gamma \gets \mathsf{Taint}_{S,L}(\phi, T, \Sigma, Z, P)$: Takes as input an initial assignment $\phi$ of some taint sets $T = \{\tau_i\}$ to 
+    a set of sources $\Sigma = \{\sigma_i\}$, and an input program $P$. Outputs an 
+    assignment $\Gamma$ of statements $s_i$ in the program $P$ to taint sets $\rho_i$ according to the taint propagation semantics $S$ for programming
+    language $L$.
+\end{itemize}
 
@@
 
-\bibliographystyle{alpha}
-\bibliography{refs}
+{\normalsize \bibliographystyle{acm}
+\bibliography{refs}}
 

### 2018-12-13 paper/paper.tex
latest

@@
 
+\subsection{Related Work}
+
 \section{Preliminaries}
@@
     a set of sources $\Sigma = \{\sigma_i\}$, and an input program $P$. Outputs an 
-    assignment $\Gamma$ of statements $s_i$ in the program $P$ to taint sets $\rho_i$ according to the taint propagation semantics $S$ for programming
+    assignment $\Gamma$ of statements $s_i$ in the program $P$ to taint sets $\tau_i \in T$ according to the taint propagation semantics $S$ for programming
     language $L$.

### 2018-12-13 paper/paper.tex
Inference rules and instrumentation sections

@@
 \newcommand{\cross}{\otimes{}}
+
+\usepackage[T1]{fontenc}
 
@@
 
-Description of instrumentation algorithm.
+	% Take C/C++ project and turn it into our IR.
+
+	In the instrumentation phase, we take an entire software project, and augment it with the other inputs (like annotations) to prepare it for the static analysis we will perform.
+Ultimately, our goal is to make assertions about the taintsets of program variables at the sinks.
+But first, we will individually instrument each version of the software project.
+
+	First we transform the code from its existing state into the language whose semantics we described earlier.
+We also rename the variables in the second version of the program so their names are disjoint from the variable names in program one.
+Then, in both programs, we replace the legitimate sources of entropy with labelled constants.
+We propogate these labelled constants through the program to propagate the taint of the entropy.
+In the process of doing so, it is possible that a value is tainted by more than one source (which is why we are tracking the taintset).
+We replace statements that taint a value with two or more sources ($s_1, s_2, \ldots, s_n$) with one of two uninterpreted functions:
+
+	\begin{enumerate}
+		\item $\textsf{preserving}(s_1, s_2, \ldots, s_n)$ if the operation preserves entropy (+, XOR, etc.)
+		\item $\textsf{non-preserving}(s_1, s_2, \ldots, s_n)$ if the operation does not preserve entropy (<<, >>, etc.)
+	\end{enumerate}
+
+	These uninterpreted functions are inferred to be pure functions over their variables, so they capture the notion of the taintset at that point.
+This allows us to assert their equality in the program we pass to CPAChecker.
+
+	Next, we will collect information from the two versions of the program as we transform them that will guide our future inference rules.
+Namely, we will perform taint analysis on sources to populate the environment $\Gamma$ which marks statements involving tainted variables.
+$\Gamma \vdash S_1$ if $S_1$ references a variable that is tainted by a source with sufficiently high entropy.
+
+	% Notice that if $x, y$ are both sufficiently entropic values then $x + y$ is also sufficiently entropic because the audition does not remove the entropy.
+Therefore addition is replaced with the $\textsf{preserving}$ function call.
+Contrast this with the left-shift, for example, which will remove entropy from the value.
+
+	% The next step involves collecting static information about the program that will help us apply our inference rules. Namely, we will populate the environment $\Gamma$ such that
+
+  %   \item Replace sources with labelled constants
+  %   \item For values that are tainted by more than one source (for example $S_1 + S_2$) replace with one of two uninterpreted functions over the sources: \begin{enumerate}
+		% \item $\textsf{preserving}(s_1, s_2, \ldots, s_n)$. Ex: $+, \oplus$
+		% \item $\textsf{non-preserving}(s_1, s_2, \ldots, s_n)$. Ex: $<<, >>$
+  %   \end{enumerate}
+  %   \item Perform taint analysis on sources to generate environment $\Gamma$ which marks statements involving tainted variables.
 
@@
 
-Description of product program inference rules approach.
+With the instrumented programs, we can construct a product program that is semantically equivalent to the sequential composition of the two programs.
+However, we would like this product program to be more amenable to reason about for the static analysis tool.
+To this end, we would like to synchronize the control flow of the two programs as much as possible.
+
+We present inference rules for composing statements in the two programs $S_1 \cross S_1$ into one program that is semantically equivalent to $S_1\ ;\ S_2$.
+The most basic inference rule simply involves sequentially composing two atomic statements, which do not involve any control flow to be syncrhonized.
+Also, because the variables in the two versions of the program have been renamed to be disjoint, we can introduce an inference rule for the commutativity of two statements.
 
@@
 
+Next, we have the inference rule for synchronizing an if-statement with another statement.
+If $S_1$ is the taken branch, $S_2$ is the not-taken branch, and $S$ is the statement after the entire if-else-statement, then we would like to inject $S$ separately into each branch.
+This simple transformation is clearly equivalent to the sequential composition of the if-statement and $S$.
+The benefit of this transformation, however, is that the predicate abstraction tool has more information about the trace taken with which it can reason about product program.
+
+Finally, we have the inference rule for synchronizing two while loops.
+The intuition here is that we would like the two loops to operate in lock-step as long as possible.
+To maintain the equivalence with the sequential composition, we also run the remaining iterations in the leftover loop after the other loop's condition no longer holds.
+
 \begin{figure}
@@
 			} \\ \\ \\
-			\infer[]{while(p_1)\ S_1 \cross while(p_2)\ S_2 \leadsto P_0\ ;\ P_1\ ;\ P_2}{
+			\infer[]{while(p_1)\ S_1 \cross while(p_2)\ S_2 \leadsto P_0\ \cross\ P_1\ ;\ P_2}{
 				\text{$P_0 = while(p_1 \land p_2)\ S_1\ ;\ S_2$} \\
@@
 
+Although these inference rules are semantically equivalent to the sequential composition, and clearly provide benefit to CPAChecker, this proving strength comes at a cost.
+Namely, the if statements and the while loops lead to an exponential blowup in the size of the product program being reasoned about as a result of the copied statements.
+
 \subsection{Taint-augmented Inference Rules}
 
-Why the naive approach doesn't work (intractability). Introduce heuristic-based approach.
+We conjecture that the exponential blowup introduced by the naive inference rules is avoidable in some cases.
+In particular, our key insight is that we don't need the extra proving power provided by the synchronization in parts of the program that are unrelated to entropy.
+For these program statements, we can resort to the sequential composition.
+For example, consider a large software program: only a small portion of the code base will be related to the entropy.
+Therefore, we should concentrate the bulk of our analysis on this small portion of the code.
+
+We define these ``unrelated'' statements to be the ones that are not tainted by any legitimate entropy source.
+This is reasonable because our ultimate assertions we would like to prove will only be over the tainted variables.
+Thus, we can use the $\Gamma$ environment we constructed in the instrumentation step to augment our inference rules.
+We say that $\Gamma \vdash S$ if some variable referenced in $S$ is tainted by at least one legitimate source of entropy.
+Similarly, $\Gamma \nvdash S$ if no variable in the statement is tainted by any legitimate source of entropy.
+We introduce an optimization inference rule that does not lead to any exponential blowup for statements that are not entailed by $\Gamma$.
 
@@
 
+Then, we can make the corresponsding changes to our naive inference rules to recognize the $\Gamma$ environment we've introduced.
+This gives us the following set of augmented inference rules: % TODO: reference to figure
+
 \begin{figure}
@@
 			} \\ \\ \\
+			\infer[]{\Gamma \vdash S_1 \cross S_2 \leadsto S_1\ ;\ S_2}{
+				\Gamma \nvdash S_1 \\
+				\Gamma \nvdash S_2
+			} \\ \\ \\
 			\infer[]{\Gamma \vdash if(p)\ then\ S_1\ else\ S_2 \cross S \leadsto P}{
@@
 				\text{$P_1 = while(p_1)\ S_1$} \\
-				\text{$P_2 = while(p_2)\ S_2$}
+				\text{$P_2 = while(p_2)\ S_2$} \\
+				\text{$P = if(p)\ then\ S_1'\ else\ S_2'$}
 			}
@@
 \end{figure}
+
+% TODO: Notice that this optimization is only a heuristic. % (taint analysis isn't sufficient given the code snippet Isil mentioned).
 
@@
 
+% TODO
+
 \section{Conclusion}
@@
 
-\section{Future Work}
+% \section{Future Work}
 
-Verify the following conjectures. Prove that a heuristic-based approach is no less powerful than a full product program approach.
-Run this tool on projects in the wild to determine how widespread entropy-misuse bugs are among software projects ``in the wild.''
+% Verify the following conjectures. Prove that a heuristic-based approach is no less powerful than a full product program approach.
+% Run this tool on projects in the wild to determine how widespread entropy-misuse bugs are among software projects ``in the wild.''
 
-\section{Considered Approaches}
+% \section{Considered Approaches}
 
-\subsection{Taint Analysis}
-Static code analysis has a history of identifying security vulnerabilities at a source code level.
+% \subsection{Taint Analysis}
+% Static code analysis has a history of identifying security vulnerabilities at a source code level.
 Some examples include SQL injection, cross-site scripting exploits, and buffer overflow attacks.
@@
 We claim that standard static code analysis techniques developed thus far are insufficient for the analysis we would like to perform.
-The sources of entropy can be tainted standard taint analysis techniques can be used to approximate how the taint propagates.
+% The sources of entropy can be tainted standard taint analysis techniques can be used to approximate how the taint propagates.
 But this approximation may be too coarse to provide useful information to the user.
@@
 
-\subsection{Differential Taint Analysis}
+% \subsection{Differential Taint Analysis}
 
-Since we want a stronger notion of taint analysis than a naive approach, we intend to introduce and formalize the concept of ``differential taint analysis." Differential
-taint analysis, in the ideal, will give us more information than simple taint analysis by leveraging 
-the original version of the program as a source of ground truth. This additional information could include reducing programmer annotation work or reduce false positives between versions
-(since taint analysis overapproximates data flow). By viewing the original version of the program as bug-free, we reduce the number of code audits required:
-only the original version of the code would need to be audited, and future versions could rely on our tool to automatically detect bugs that were introduced by refactors.
+% Since we want a stronger notion of taint analysis than a naive approach, we intend to introduce and formalize the concept of ``differential taint analysis." Differential
+% taint analysis, in the ideal, will give us more information than simple taint analysis by leveraging 
+% the original version of the program as a source of ground truth. This additional information could include reducing programmer annotation work or reduce false positives between versions
+% (since taint analysis overapproximates data flow). By viewing the original version of the program as bug-free, we reduce the number of code audits required:
+% only the original version of the code would need to be audited, and future versions could rely on our tool to automatically detect bugs that were introduced by refactors.
 
-\subsection{Product Programs}
+% \subsection{Product Programs}
 
-One technique that we have considered toward the construction of differential taint analysis is the notion of a ``product program.'' In the most general sense, the product $P_1 \times P_2$ of two programs $P_1$ and $P_2$ is used to verify
-relations between the programs, such as equivalence. Product programs have also been used to analyze different runs of the same program. 
-The product program $P_1 \times P_2$ is semantically equivalent to the sequential composition $P_1; P_2$, but such that we can prove useful safety properties of $P_1 \times P_2$ that would be difficult to prove with standard techniques on $P_1$, $P_2$, or $P_1; P_2$.
+% One technique that we have considered toward the construction of differential taint analysis is the notion of a ``product program.'' In the most general sense, the product $P_1 \times P_2$ of two programs $P_1$ and $P_2$ is used to verify
+% relations between the programs, such as equivalence. Product programs have also been used to analyze different runs of the same program. 
+% The product program $P_1 \times P_2$ is semantically equivalent to the sequential composition $P_1; P_2$, but such that we can prove useful safety properties of $P_1 \times P_2$ that would be difficult to prove with standard techniques on $P_1$, $P_2$, or $P_1; P_2$.
 
-In our setting, we are not so much interested in proving safety properties of a program $P$ but instead wish to prove that cryptographic values rely on sufficiently many bits of entropy when they are used via taint analysis,
-which is why we cannot simply leverage existing off-the-shelf product program analysis techniques.
+% In our setting, we are not so much interested in proving safety properties of a program $P$ but instead wish to prove that cryptographic values rely on sufficiently many bits of entropy when they are used via taint analysis,
+% which is why we cannot simply leverage existing off-the-shelf product program analysis techniques.
 
-\subsubsection{Example}
+% \subsubsection{Example}
 
-Consider the following two programs $P_1$ and $P_2$ that are equivalent from a cryptographic point of view, but might have small feature changes.
+% Consider the following two programs $P_1$ and $P_2$ that are equivalent from a cryptographic point of view, but might have small feature changes.
 
-\begin{algorithmic}
-\Function{$P_1$}{}
-\State $s \gets$ entropy
-\State $\vdots$
-\If {$p$}
-\State $c \gets$ AES(s)
-\EndIf
-\State $\vdots$
-\EndFunction
-\\
-\Function{$P_2$}{}
-\State $s \gets$ entropy
-\State $\vdots$
-\If {$p$}
-\State $c \gets$ AES(s)
-\EndIf
-\State $\vdots$
-\EndFunction
-\end{algorithmic}
+% \begin{algorithmic}
+% \Function{$P_1$}{}
+% \State $s \gets$ entropy
+% \State $\vdots$
+% \If {$p$}
+% \State $c \gets$ AES(s)
+% \EndIf
+% \State $\vdots$
+% \EndFunction
+% \\
+% \Function{$P_2$}{}
+% \State $s \gets$ entropy
+% \State $\vdots$
+% \If {$p$}
+% \State $c \gets$ AES(s)
+% \EndIf
+% \State $\vdots$
+% \EndFunction
+% \end{algorithmic}
 
-where $p$ is some complex predicate that is difficult to reason about, but does not change between $P_1$ and $P_2$ (and no values that $p$ depends on change, either).
+% where $p$ is some complex predicate that is difficult to reason about, but does not change between $P_1$ and $P_2$ (and no values that $p$ depends on change, either).
 Then, a safety property about the entropy of the cryptographic values in $c$ would hold in both programs or in neither program.
 
-\subsection{Verification Modulo Versions}
-Verification modulo versions is an idea quite similar to product programs. VMV attempts to identify abstract regressions and relative correctness between two versions of a program.
-In VMV, the static analyzer is treated as opaque in an effort to cut down the number of alarms by inferring assumptions made by the original program (assumed to be correct).
+% \subsection{Verification Modulo Versions}
+% Verification modulo versions is an idea quite similar to product programs. VMV attempts to identify abstract regressions and relative correctness between two versions of a program.
+% In VMV, the static analyzer is treated as opaque in an effort to cut down the number of alarms by inferring assumptions made by the original program (assumed to be correct).
 This is especially appealing in the security software development cycle because there are documented cases of entropy bugs being introduced in new versions of programs (for example, OpenSSL and FreeBSD).
@@
 
-\subsection{Relational Verification}
-Another existing technique is relational verification, in which relational properties between two programs are used to prove $k$-safety properties.
+% \subsection{Relational Verification}
+% Another existing technique is relational verification, in which relational properties between two programs are used to prove $k$-safety properties.
 A $k$-safety property is a safety property of a program that requires
-reasoning about the relationships between $k$ different runs of a program (or $k$ programs). For example, transitivity is a $k$-safety property.
-In our case, we would want to show the equivalence of entropy assignment and the equivalence of the entropy propagation.
+% reasoning about the relationships between $k$ different runs of a program (or $k$ programs). For example, transitivity is a $k$-safety property.
+% In our case, we would want to show the equivalence of entropy assignment and the equivalence of the entropy propagation.
 Relational verification utilitzes product programs; RV constructs a product program between versions of the program that captures the semantics of executing them sequentially, but that lends itself to standard verification techniques.
 
-\subsection{Differential Assertion Checking}
+% \subsection{Differential Assertion Checking}
 
-The last approach that we have considered is differential assertion checking. Differential assertion checking seeks to (statically) prove the relative correctness between two similar programs,
-with a significantly lower cost than ensuring absolute correctness. That is, DAC allows is to answer if there are conditions under which a program $P_1$ passes an assertion check but
-a program $P_2$ fails. The techniques utilized current DAC work may be of interest, but it seems they rely on an a priori mapping between the two programs, which is something that
-relational verification work seeks to answer. Thus DAC will be considered further after the idea of product programs is exhausted.
+% The last approach that we have considered is differential assertion checking. Differential assertion checking seeks to (statically) prove the relative correctness between two similar programs,
+% with a significantly lower cost than ensuring absolute correctness. That is, DAC allows is to answer if there are conditions under which a program $P_1$ passes an assertion check but
+% a program $P_2$ fails. The techniques utilized current DAC work may be of interest, but it seems they rely on an a priori mapping between the two programs, which is something that
+% relational verification work seeks to answer. Thus DAC will be considered further after the idea of product programs is exhausted.
 
-% How is vmv related to relational verification?
+% % How is vmv related to relational verification?
 
-% Standard taint analysis will under/overapproximate the taint, so we may not be able to use it to directly prove our safety property on the target program.
+% % Standard taint analysis will under/overapproximate the taint, so we may not be able to use it to directly prove our safety property on the target program.
 Thus, we can leverage the approach verification modulo versions to prove relational verification.
 
-% Relational verification aims to solve this problem by proving relational properties between two programs.
+% % Relational verification aims to solve this problem by proving relational properties between two programs.
 However, the naive relational verification approach of verifying each program individually is infeasible given arbitrarily difficult-to-reason-about functions even if those functions do not change between versions of the program.
@@
 
-% Finally, even with a sufficiently precise taint analysis, it is unclear how to determine the ``correct'' set of sources a cryptographic value should rely on at any point in the program.
+% % Finally, even with a sufficiently precise taint analysis, it is unclear how to determine the ``correct'' set of sources a cryptographic value should rely on at any point in the program.
 For example, some programs will use weak entropy on system startup, and increase the entropy in those values, later on.
@@
 
-\section{Our Approach}
-Given the two programs in the example, we propose generating a product program similar to the following:
+% \section{Our Approach}
+% Given the two programs in the example, we propose generating a product program similar to the following:
 
-\begin{algorithmic}
-\Function{$P_1 \times P_2$}{}
-\State $S_1 \gets$ entropy
-\State $S_2 \gets$ entropy
-\State $\vdots$
-\If {f()}
-\State c$_1 \gets$ AES($s_1$)
-\State $c_2 \gets$ AES($s_2$)
-\EndIf
-\State $\vdots$
-\EndFunction
-\end{algorithmic}
+% \begin{algorithmic}
+% \Function{$P_1 \times P_2$}{}
+% \State $S_1 \gets$ entropy
+% \State $S_2 \gets$ entropy
+% \State $\vdots$
+% \If {f()}
+% \State c$_1 \gets$ AES($s_1$)
+% \State $c_2 \gets$ AES($s_2$)
+% \EndIf
+% \State $\vdots$
+% \EndFunction
+% \end{algorithmic}
 
-Note that, during the construction of the product program, we need to be able to merge the if statements in the two programs.
+% Note that, during the construction of the product program, we need to be able to merge the if statements in the two programs.
 
-We would like to demonstrate set equality between the taint for $s_1, s_2$ when they are used in the if statement, which amounts to demonstrating the equivalence of the taint propagation through the program.
+% We would like to demonstrate set equality between the taint for $s_1, s_2$ when they are used in the if statement, which amounts to demonstrating the equivalence of the taint propagation through the program.
 We would also like to demonstrate that we assign the output of a sufficiently entropic value to a variable in $P_1$ if and only if we also assign it that sufficiently entropic value in $P_2$.
 
-Proving these two properties (set-equality and equivalence of assignments) allows us to claim that $P_1$ has the safety property if and only if $P_2$ has the safety property.
+% Proving these two properties (set-equality and equivalence of assignments) allows us to claim that $P_1$ has the safety property if and only if $P_2$ has the safety property.
 
-\section{Real World Applications}
+% \section{Real World Applications}
 
-A verification tool like this could be especially useful for a maintainer of a project that relies on cryptographic values.
+% A verification tool like this could be especially useful for a maintainer of a project that relies on cryptographic values.
 For example, if she audits her code once to confirm that it uses entropy properly, she can use differential taint analysis on future commits to confirm that the new code does not introduce this class of bugs.
@@
 
-Our tool should be able to detect the infamous Debian OpenSSL entropy bug.
+% Our tool should be able to detect the infamous Debian OpenSSL entropy bug.
 We will also test our tool on other libraries that rely on entropy, such as Amazon's signal2noise, LibreSSL, GnuTLS, and GnuPG,
-which are not known to contain entropy bugs (and likely do not), but we can test on versions in which we introduce our own bugs.
+% which are not known to contain entropy bugs (and likely do not), but we can test on versions in which we introduce our own bugs.
 
-\section{Research Hypotheses}
+% \section{Research Hypotheses}
 
-These are the principal hypotheses we would like to test:
+% These are the principal hypotheses we would like to test:
 
-\begin{enumerate}
-	\item An automated tool can detect entropy bugs in real-world programs.
-	\item Entropy is insufficiently propagated in programs that rely on cryptography, or entropy propagation follows nontrivial code paths (due to error handling or other control flow).
-	\item Multiple versions of the same program can make static analysis for this domain more effective by lowering costs (computational and programmer) or providing more fine-grained information.
-\end{enumerate}
+% \begin{enumerate}
+% 	\item An automated tool can detect entropy bugs in real-world programs.
+% 	\item Entropy is insufficiently propagated in programs that rely on cryptography, or entropy propagation follows nontrivial code paths (due to error handling or other control flow).
+% 	\item Multiple versions of the same program can make static analysis for this domain more effective by lowering costs (computational and programmer) or providing more fine-grained information.
+% \end{enumerate}
 
-\section{Future Work}
+% \section{Future Work}
 
-As bad data flow can be a source of security bugs other than simply (lack of) entropy propagation (for example, 
-address disclosure and use-after-free bugs in browsers), we believe that our tool could also be applied to these settings to help prevent
-programmers from introducing security vulnerabilities into their projects.
+% As bad data flow can be a source of security bugs other than simply (lack of) entropy propagation (for example, 
+% address disclosure and use-after-free bugs in browsers), we believe that our tool could also be applied to these settings to help prevent
+% programmers from introducing security vulnerabilities into their projects.
 
-\section{Links}
+% \section{Links}
 
-\begin{enumerate}
-	\item Debian/OpenSSL Bug 
-		\begin{enumerate}
-			\item \url{https://www.schneier.com/blog/archives/2008/05/random_number_b.html}
-			\item \url{https://research.swtch.com/openssl}
-			\item \url{https://freedom-to-tinker.com/2013/09/20/software-transparency-debian-openssl-bug/}
-			\item \url{https://www.cs.umd.edu/class/fall2017/cmsc818O/papers/private-keys-public.pdf}
-		\end{enumerate}
-	\item Data flow
-		\begin{enumerate}
-			\item \url{https://en.wikipedia.org/wiki/Data-flow_analysis}
-			\item \url{https://www.seas.harvard.edu/courses/cs252/2011sp/slides/Lec02-Dataflow.pdf}
-		\end{enumerate}
-	\item Static Program Analysis
-		\begin{enumerate}
-            \item \url{https://cs.au.dk/~amoeller/spa/spa.pdf}
-            \item \url{https://ieeexplore.ieee.org/stamp/stamp.jsp?arnumber=6859783}
-		\end{enumerate}
-	\item Relational Verification:
-		\begin{enumerate}
-			\item \url{https://dl.acm.org/citation.cfm?id=2021319}
-			\item \url{https://ac.els-cdn.com/S235222081630044X/1-s2.0-S235222081630044X-main.pdf?_tid=076a0492-9cee-4995-9710-bcb3c64b98e0&acdnat=1539815890_178849b4f14af3751e9acb03b238db4d}
-			\item \url{https://www.microsoft.com/en-us/research/publication/differential-assertion-checking/}
-			\item \url{https://www.microsoft.com/en-us/research/wp-content/uploads/2014/06/paper-1.pdf}
-			\item \url{https://www.cs.utexas.edu/~isil/pldi16-chl.pdf}
-		\end{enumerate}
-	\item Projects to analyze 
-		\begin{enumerate}
-			\item OpenPGP
-			\item BouncyCastle
-			\item OpenSSL
-			\item GnuPGP
-			\item F\# SSL project with proof of correctness
-			\item NQSBTLS
-			\item Amazon's s2n (signal to noise)
-		\end{enumerate}
-\end{enumerate}
+% \begin{enumerate}
+% 	\item Debian/OpenSSL Bug 
+% 		\begin{enumerate}
+% 			\item \url{https://www.schneier.com/blog/archives/2008/05/random_number_b.html}
+% 			\item \url{https://research.swtch.com/openssl}
+% 			\item \url{https://freedom-to-tinker.com/2013/09/20/software-transparency-debian-openssl-bug/}
+% 			\item \url{https://www.cs.umd.edu/class/fall2017/cmsc818O/papers/private-keys-public.pdf}
+% 		\end{enumerate}
+% 	\item Data flow
+% 		\begin{enumerate}
+% 			\item \url{https://en.wikipedia.org/wiki/Data-flow_analysis}
+% 			\item \url{https://www.seas.harvard.edu/courses/cs252/2011sp/slides/Lec02-Dataflow.pdf}
+% 		\end{enumerate}
+% 	\item Static Program Analysis
+% 		\begin{enumerate}
+%             \item \url{https://cs.au.dk/~amoeller/spa/spa.pdf}
+%             \item \url{https://ieeexplore.ieee.org/stamp/stamp.jsp?arnumber=6859783}
+% 		\end{enumerate}
+% 	\item Relational Verification:
+% 		\begin{enumerate}
+% 			\item \url{https://dl.acm.org/citation.cfm?id=2021319}
+% 			\item \url{https://ac.els-cdn.com/S235222081630044X/1-s2.0-S235222081630044X-main.pdf?_tid=076a0492-9cee-4995-9710-bcb3c64b98e0&acdnat=1539815890_178849b4f14af3751e9acb03b238db4d}
+% 			\item \url{https://www.microsoft.com/en-us/research/publication/differential-assertion-checking/}
+% 			\item \url{https://www.microsoft.com/en-us/research/wp-content/uploads/2014/06/paper-1.pdf}
+% 			\item \url{https://www.cs.utexas.edu/~isil/pldi16-chl.pdf}
+% 		\end{enumerate}
+% 	\item Projects to analyze 
+% 		\begin{enumerate}
+% 			\item OpenPGP
+% 			\item BouncyCastle
+% 			\item OpenSSL
+% 			\item GnuPGP
+% 			\item F\# SSL project with proof of correctness
+% 			\item NQSBTLS
+% 			\item Amazon's s2n (signal to noise)
+% 		\end{enumerate}
+% \end{enumerate}
 
-{\normalsize \bibliographystyle{acm}
-\bibliography{refs}}
+% {\normalsize \bibliographystyle{acm}
+% \bibliography{refs}}
 
-% We have two motivating examples. First, we would like our tool to be able to detect the issue with the Debian/OpenSSL pseudo-random number generator that was exposed in 2008. 
-% Second, we would like to identify potential vulnerabilities in current cryptocurrency wallet code as many cryptocurrency protocols rely on DSA.
+% % We have two motivating examples. First, we would like our tool to be able to detect the issue with the Debian/OpenSSL pseudo-random number generator that was exposed in 2008. 
+% % Second, we would like to identify potential vulnerabilities in current cryptocurrency wallet code as many cryptocurrency protocols rely on DSA.
 
-% \subsection{Goals}
-% We plan to use data dependency tools to determine how entropic inputs in a given program are used by various cryptographic algorithms. 
-% That will allow us to identify if and when entropy is too low or is misused. We plan to either produce
-% this as a code integration tool for developers to use as part of a compiler toolchain, or use this tool to analyze a large number of codebases
-% found ``in the wild,'' such as those written by amateurs.
+% % \subsection{Goals}
+% % We plan to use data dependency tools to determine how entropic inputs in a given program are used by various cryptographic algorithms. 
+% % That will allow us to identify if and when entropy is too low or is misused. We plan to either produce
+% % this as a code integration tool for developers to use as part of a compiler toolchain, or use this tool to analyze a large number of codebases
+% % found ``in the wild,'' such as those written by amateurs.
 
-% Our tool will seek to generate a human-checkable dependency graph from source code
-% using taint analysis on functions and function inputs manually specified via annotation by the developer.
-% This graph will allow the developer to manually verify that entropy is used properly; a stretch goal 
-% of ours would be to automate the identification of a subset of known misuses of entropy.
+% % Our tool will seek to generate a human-checkable dependency graph from source code
+% % using taint analysis on functions and function inputs manually specified via annotation by the developer.
+% % This graph will allow the developer to manually verify that entropy is used properly; a stretch goal 
+% % of ours would be to automate the identification of a subset of known misuses of entropy.
 

### 2018-12-13 paper/paper.tex

--- 2018-12-13 paper/paper.tex
@@
 \usepackage{algorithm}
+\usepackage{algorithmicx}
 \usepackage{algpseudocode}
@@
 %make title bold and 14 pt font (Latex default is non-bold, 16 pt)
-\title{\Large \bf New Static Analysis Techniques to Detect Entropy Failure Vulnerabilities}
+\title{\Large \bf New Static Analysis Techniques to Detect Entropy Failure Vulnerabilities in Modern Software Projects}
 
@@
 
+Differential assertion checking, verification modulo versions, relational verification using product programs, secure information flow as a safety problem
+
 \section{Preliminaries}
 
-\subsection{Toy Language}
+\subsection{Demonstrative Language}
 
@@
 
-Our results can be extended in a straightforward manner to an industrial standard language such as C.
+Our results can be extended in a straightforward manner to an industrial language such as C, and we do so for our evaluation (Section \ref{sec:eval}).
 
@@
 As opposed to a safety property of a program (for example, ``variable $x$ is always positive'' or ``pointer $p$ is never null''), which requires
-the absence of errors in a single program trace, a $k$-safety property of a program requires the absence of erroneous interactions between $k$
+the absence of errors in a single program trace \cite{alpern1987recognizing}, a $k$-safety property of a program requires the absence of erroneous interactions between $k$
 traces of the same program \cite{sousa2016cartesian}. A program property such as symmetry is a $2$-safety property.
@@
 Predicate abstraction is a specialized form of abstract interpretation that can be used for checking $1$-safety properties of programs
-\cite{flanagan2002predicate}.
+\cite{flanagan2002predicate}. Abstract interpretation is a static analysis technique for programs that allows a verifier to, in some cases, 
+automatically infer properties such as loop invariants. Formally we have:
+
+\begin{itemize}
+    \item $b \gets \mathsf{Verify}_{L}(\pi, P)$: Takes as input a $1$-safety property $\pi$ and a program $P$ and outputs an accepting bit $\top$ if
+    $P$ satisfies the safety property, and otherwise outputs a rejecting bit $\bot$ or fails to terminate otherwise.
+\end{itemize}
+
+CPAchecker \cite{beyer2011cpachecker} is an example of a state-of-the-art predicate abstraction verifier for the C programming language. In this work
+we are interested in the safety property of equality. We note that due to inherent hardness results such as NP-hardness and undecidability, predicate
+abstraction verifiers are based on heuristic methods and are not guaranteed to run in polynomial time, or even terminate.
 
@@
 
+The product $P_1 \times P_2$ of two programs $P_1$ and $P_2$ is not an entirely well-defined concept. In the most general sense, the product $P_1 \times P_2$ of 
+two programs $P_1$ and $P_2$ is used to verify
+relations between the programs, such as equivalence \cite{barthe2011relational}. Product programs have also been used to analyze different runs of the same program. 
+The product program $P_1 \times P_2$ is semantically equivalent to the sequential composition $P_1; P_2$, but such that we can prove useful 
+safety properties of $P_1 \times P_2$ that would be difficult to 
+prove with standard techniques on $P_1$, $P_2$, or $P_1; P_2$.
+
+For our purposes we simply require that $P_1 \times P_2$ is semantically
+equivalent to the sequential composition $P_1\ ;\ P_2$ of the two programs. Our goal in using the concept of product programs is to aid the predicate abstraction
+verifier.
+
 \section{Our Approach}
 
-Want to use differential view since we can get more out of this. Unsound taint analysis example.
+Our main approach to this problem is to take two versions $P_1$ and $P_2$ of a program and prove $P_2$ \emph{relative} to $P_1$. We have several
+motivations for this approach:
+
+\begin{itemize}
+    \item Taint analysis is an (over)approximate solution.
+    \item Modern software projects often use version control history, and we can leverage that to give a static verifier more power (heuristically speaking).
+\end{itemize}
+
+Let us address the first point.
+An initial approach to this problem
+might consist of using off-the-shelf taint analysis (we would taint high-entropy sources and confirm that
+cryptographic functions receive high-entropy tainted inputs). 
+However, taint analysis merely provides an overapproximation to the actual entropy that is
+input to a cryptographic algorithm. Consider the program in Figure
+\ref{fig:prog1}. The variable $k$ is tainted with high-entropy, but will not contain high-entropy if $f()$ ever evaluates to false.
+
+\begin{figure}
+    \label{fig:prog1}
+    \caption{Taint analysis overapproximation}
+    \begin{center}
+    \begin{algorithmic}
+
+        \If {$f()$}
+        \State $k \gets$ \textsf{read(128, ``/dev/urandom'')}
+        \Else
+        \State $k \gets$ \textsf{getpid()}
+        \EndIf
+        \State $\mathsf{ctxt} \gets $ \textsf{AES}$(k,\mathsf{ptxt})$
+    \end{algorithmic}
+\end{center}
+\end{figure}
+
+Now we consider the second point. Since software developers often have access to multiple revisions of a single program, we wish to use this additional information
+to do better than simple taint analysis on a single version. This allows for stronger security guarantees at lower cost: a software project can invest
+in manually auditing an initial version of its security-critical software, and then use our methods to prove that subsequent versions do not contain entropy
+failures provided that this initial version does not (i.e., a \emph{relative} or \emph{differential} notion of security). Additionally, since we expect
+program revisions (particularly those touching security-critical components) to be infrequent and small, we can reduce the scope of the verifier's work.
+
+\subsection{A First Attempt}
+
+Given that we wish to attain this differential notion of security, we first address a naive attempt at solving this problem using taint analysis.
+Consider the program in Figure \ref{fig:prog2} compared to the program in Figure \ref{fig:prog1}. The taint sets of $k$ in both versions of the program
+are equal, however, simply comparing for equality here is \emph{unsound} with respect to our differential notion of security, since for all program traces
+in which the first program is secure, all similar program traces in the second program are insecure.
+
+\begin{figure}
+    \label{fig:prog2}
+    \caption{Taint analysis unsoundness}
+    \begin{center}
+    \begin{algorithmic}
+        \If {$f()$}
+        \State $k \gets$ \textsf{getpid()}
+        \Else
+        \State $k \gets$ \textsf{read(128, ``/dev/urandom'')}
+        \EndIf
+        \State $\mathsf{ctxt} \gets $ \textsf{AES}$(k,\mathsf{ptxt})$
+    \end{algorithmic}
+\end{center}
+\end{figure}
 
@@
 
+\subsection{Overview}
+
+Our solution proceeds in three primary stages:
+
+\begin{itemize}
+    \item \textbf{Instrumentation:} In this step we reduce our taint analysis problem for each program version $P_1$ and $P_2$ to one of $2$-safety.
+    Additionally, we run taint analysis on each program to get outputs $\Gamma_1$ and $\Gamma_2$, which we use in the next step.
+    \item \textbf{Program syncronization:} Using our environment $\Gamma = \Gamma_1 \cup \Gamma_2$, our instrumented programs $P'_1$ and $P'_2$ from 
+    the previous step, we construct the product program $P'_1 \times P'_2$. This converts the $2$-safety problem to one of $1$-safety.
+    \item \textbf{Verification:} Finally, we pass $P'_1 \times P'_2$ to an off-the-shelf verifier to check the $1$-safety property to
+    prove security.
+\end{itemize}
+
+We describe these in more formal detail in the coming sections.
+
 \subsection{Instrumenting Programs for Predicate Abstraction}
+
+\textbf{Instrumentation:} Given two program versions $P_1$ and $P_2$, our sources $\Sigma$, and a taint analysis algorithm $\mathsf{Taint}_{S,L}$
+     we reduce the taint analysis problem to one of $1$-safety.
+    We do this for each $P_k$ by assigning constants $c$ to all sources $\sigma \in \Sigma$, and then where we would propagate taint using the rules $S$ for language $L$
+    in statements $s_i$ which is a function of program variables $v_i$ and only of ``entropy preserving" operations (described below) 
+    we replace with an uninterpreted function $f(v_1, \ldots, v_j)$. For statements $s'_i$ which are a
+    function of any tainted program variables $w_j$ and contain any ``non-entropy preserving operations'' (any operations that are not entropy preserving), we replace
+    with an uninterpreted function $g(w_1, \ldots, w_j)$. We output these programs as $P'_1$ and $P'_2$.
 
@@
 \section{Evaluation}
+\label{sec:eval}
 
@@
 
-\section{Our Approach}
-Given the two programs in the example, we propose generating a product program similar to the following:
-
-\begin{algorithmic}
-\Function{$P_1 \times P_2$}{}
-\State $S_1 \gets$ entropy
-\State $S_2 \gets$ entropy
-\State $\vdots$
-\If {f()}
-\State c$_1 \gets$ AES($s_1$)
-\State $c_2 \gets$ AES($s_2$)
-\EndIf
-\State $\vdots$
-\EndFunction
-\end{algorithmic}
-
-Note that, during the construction of the product program, we need to be able to merge the if statements in the two programs.
-
-We would like to demonstrate set equality between the taint for $s_1, s_2$ when they are used in the if statement, which amounts to demonstrating the equivalence of the taint propagation through the program.
-We would also like to demonstrate that we assign the output of a sufficiently entropic value to a variable in $P_1$ if and only if we also assign it that sufficiently entropic value in $P_2$.
-
-Proving these two properties (set-equality and equivalence of assignments) allows us to claim that $P_1$ has the safety property if and only if $P_2$ has the safety property.
-
-\section{Real World Applications}
-
-A verification tool like this could be especially useful for a maintainer of a project that relies on cryptographic values.
-For example, if she audits her code once to confirm that it uses entropy properly, she can use differential taint analysis on future commits to confirm that the new code does not introduce this class of bugs.
-% VMV?
-
-Our tool should be able to detect the infamous Debian OpenSSL entropy bug.
-We will also test our tool on other libraries that rely on entropy, such as Amazon's signal2noise, LibreSSL, GnuTLS, and GnuPG,
-which are not known to contain entropy bugs (and likely do not), but we can test on versions in which we introduce our own bugs.
-
-\section{Research Hypotheses}
-
-These are the principal hypotheses we would like to test:
-
-\begin{enumerate}
-	\item An automated tool can detect entropy bugs in real-world programs.
-	\item Entropy is insufficiently propagated in programs that rely on cryptography, or entropy propagation follows nontrivial code paths (due to error handling or other control flow).
-	\item Multiple versions of the same program can make static analysis for this domain more effective by lowering costs (computational and programmer) or providing more fine-grained information.
-\end{enumerate}
-
-\section{Future Work}
-
-As bad data flow can be a source of security bugs other than simply (lack of) entropy propagation (for example, 
-address disclosure and use-after-free bugs in browsers), we believe that our tool could also be applied to these settings to help prevent
-programmers from introducing security vulnerabilities into their projects.
-
 \section{Links}

### 2018-12-13 paper/paper.tex

@@
 
+	% Take C/C++ project and turn it into our IR.
+
+	In the instrumentation phase, we take an entire software project, and augment it with the other inputs (like annotations) to prepare it for the static analysis we will perform.
+Ultimately, our goal is to make assertions about the taintsets of program variables at the sinks.
+But first, we will individually instrument each version of the software project.
+
+	First we transform the code from its existing state into the language whose semantics we described earlier.
+We also rename the variables in the second version of the program so their names are disjoint from the variable names in program one.
+Then, in both programs, we replace the legitimate sources of entropy with labelled constants.
+We propogate these labelled constants through the program to propagate the taint of the entropy.
+In the process of doing so, it is possible that a value is tainted by more than one source (which is why we are tracking the taintset).
+We replace statements that taint a value with two or more sources ($s_1, s_2, \ldots, s_n$) with one of two uninterpreted functions:
+
+	\begin{enumerate}
+		\item $\textsf{preserving}(s_1, s_2, \ldots, s_n)$ if the operation preserves entropy (+, XOR, etc.)
+		\item $\textsf{non-preserving}(s_1, s_2, \ldots, s_n)$ if the operation does not preserve entropy (<<, >>, etc.)
+	\end{enumerate}
+
+	These uninterpreted functions are inferred to be pure functions over their variables, so they capture the notion of the taintset at that point.
+This allows us to assert their equality in the program we pass to CPAChecker.
+
+	Next, we will collect information from the two versions of the program as we transform them that will guide our future inference rules.
+Namely, we will perform taint analysis on sources to populate the environment $\Gamma$ which marks statements involving tainted variables.
+$\Gamma \vdash S_1$ if $S_1$ references a variable that is tainted by a source with sufficiently high entropy.
+
+	% Notice that if $x, y$ are both sufficiently entropic values then $x + y$ is also sufficiently entropic because the audition does not remove the entropy.
+Therefore addition is replaced with the $\textsf{preserving}$ function call.
+Contrast this with the left-shift, for example, which will remove entropy from the value.
+
+	% The next step involves collecting static information about the program that will help us apply our inference rules. Namely, we will populate the environment $\Gamma$ such that
+
+  %   \item Replace sources with labelled constants
+  %   \item For values that are tainted by more than one source (for example $S_1 + S_2$) replace with one of two uninterpreted functions over the sources: \begin{enumerate}
+		% \item $\textsf{preserving}(s_1, s_2, \ldots, s_n)$. Ex: $+, \oplus$
+		% \item $\textsf{non-preserving}(s_1, s_2, \ldots, s_n)$. Ex: $<<, >>$
+  %   \end{enumerate}
+  %   \item Perform taint analysis on sources to generate environment $\Gamma$ which marks statements involving tainted variables.
 \textbf{Instrumentation:} Given two program versions $P_1$ and $P_2$, our sources $\Sigma$, and a taint analysis algorithm $\mathsf{Taint}_{S,L}$
@@
 
-\subsection{Differential Taint Analysis}
+% \subsection{Differential Taint Analysis}
 
-Since we want a stronger notion of taint analysis than a naive approach, we intend to introduce and formalize the concept of ``differential taint analysis." Differential
-taint analysis, in the ideal, will give us more information than simple taint analysis by leveraging 
-the original version of the program as a source of ground truth. This additional information could include reducing programmer annotation work or reduce false positives between versions
-(since taint analysis overapproximates data flow). By viewing the original version of the program as bug-free, we reduce the number of code audits required:
-only the original version of the code would need to be audited, and future versions could rely on our tool to automatically detect bugs that were introduced by refactors.
+% Since we want a stronger notion of taint analysis than a naive approach, we intend to introduce and formalize the concept of ``differential taint analysis." Differential
+% taint analysis, in the ideal, will give us more information than simple taint analysis by leveraging 
+% the original version of the program as a source of ground truth. This additional information could include reducing programmer annotation work or reduce false positives between versions
+% (since taint analysis overapproximates data flow). By viewing the original version of the program as bug-free, we reduce the number of code audits required:
+% only the original version of the code would need to be audited, and future versions could rely on our tool to automatically detect bugs that were introduced by refactors.
 
-\subsection{Product Programs}
+% \subsection{Product Programs}
 
-One technique that we have considered toward the construction of differential taint analysis is the notion of a ``product program.'' In the most general sense, the product $P_1 \times P_2$ of two programs $P_1$ and $P_2$ is used to verify
-relations between the programs, such as equivalence. Product programs have also been used to analyze different runs of the same program. 
-The product program $P_1 \times P_2$ is semantically equivalent to the sequential composition $P_1; P_2$, but such that we can prove useful safety properties of $P_1 \times P_2$ that would be difficult to prove with standard techniques on $P_1$, $P_2$, or $P_1; P_2$.
+% One technique that we have considered toward the construction of differential taint analysis is the notion of a ``product program.'' In the most general sense, the product $P_1 \times P_2$ of two programs $P_1$ and $P_2$ is used to verify
+% relations between the programs, such as equivalence. Product programs have also been used to analyze different runs of the same program. 
+% The product program $P_1 \times P_2$ is semantically equivalent to the sequential composition $P_1; P_2$, but such that we can prove useful safety properties of $P_1 \times P_2$ that would be difficult to prove with standard techniques on $P_1$, $P_2$, or $P_1; P_2$.
 
-In our setting, we are not so much interested in proving safety properties of a program $P$ but instead wish to prove that cryptographic values rely on sufficiently many bits of entropy when they are used via taint analysis,
-which is why we cannot simply leverage existing off-the-shelf product program analysis techniques.
+% In our setting, we are not so much interested in proving safety properties of a program $P$ but instead wish to prove that cryptographic values rely on sufficiently many bits of entropy when they are used via taint analysis,
+% which is why we cannot simply leverage existing off-the-shelf product program analysis techniques.
 
-\subsubsection{Example}
+% \subsubsection{Example}
 
-Consider the following two programs $P_1$ and $P_2$ that are equivalent from a cryptographic point of view, but might have small feature changes.
+% Consider the following two programs $P_1$ and $P_2$ that are equivalent from a cryptographic point of view, but might have small feature changes.
 
-\begin{algorithmic}
-\Function{$P_1$}{}
-\State $s \gets$ entropy
-\State $\vdots$
-\If {$p$}
-\State $c \gets$ AES(s)
-\EndIf
-\State $\vdots$
-\EndFunction
-\\
-\Function{$P_2$}{}
-\State $s \gets$ entropy
-\State $\vdots$
-\If {$p$}
-\State $c \gets$ AES(s)
-\EndIf
-\State $\vdots$
-\EndFunction
-\end{algorithmic}
+% \begin{algorithmic}
+% \Function{$P_1$}{}
+% \State $s \gets$ entropy
+% \State $\vdots$
+% \If {$p$}
+% \State $c \gets$ AES(s)
+% \EndIf
+% \State $\vdots$
+% \EndFunction
+% \\
+% \Function{$P_2$}{}
+% \State $s \gets$ entropy
+% \State $\vdots$
+% \If {$p$}
+% \State $c \gets$ AES(s)
+% \EndIf
+% \State $\vdots$
+% \EndFunction
+% \end{algorithmic}
 
-where $p$ is some complex predicate that is difficult to reason about, but does not change between $P_1$ and $P_2$ (and no values that $p$ depends on change, either).
+% where $p$ is some complex predicate that is difficult to reason about, but does not change between $P_1$ and $P_2$ (and no values that $p$ depends on change, either).
 Then, a safety property about the entropy of the cryptographic values in $c$ would hold in both programs or in neither program.
-
-\subsection{Verification Modulo Versions}
-Verification modulo versions is an idea quite similar to product programs. VMV attempts to identify abstract regressions and relative correctness between two versions of a program.
-In VMV, the static analyzer is treated as opaque in an effort to cut down the number of alarms by inferring assumptions made by the original program (assumed to be correct).
-This is especially appealing in the security software development cycle because there are documented cases of entropy bugs being introduced in new versions of programs (for example, OpenSSL and FreeBSD).
-This would involve under-approximating the taint of one version of the program and over-approximating the taint of another version of the program.
-Current tooling does not widely support under-approximating taint analysis.
-
-\subsection{Relational Verification}
-Another existing technique is relational verification, in which relational properties between two programs are used to prove $k$-safety properties.
-A $k$-safety property is a safety property of a program that requires
-reasoning about the relationships between $k$ different runs of a program (or $k$ programs). For example, transitivity is a $k$-safety property.
-In our case, we would want to show the equivalence of entropy assignment and the equivalence of the entropy propagation.
-Relational verification utilitzes product programs; RV constructs a product program between versions of the program that captures the semantics of executing them sequentially, but that lends itself to standard verification techniques.
-
-\subsection{Differential Assertion Checking}
-
-The last approach that we have considered is differential assertion checking. Differential assertion checking seeks to (statically) prove the relative correctness between two similar programs,
-with a significantly lower cost than ensuring absolute correctness. That is, DAC allows is to answer if there are conditions under which a program $P_1$ passes an assertion check but
-a program $P_2$ fails. The techniques utilized current DAC work may be of interest, but it seems they rely on an a priori mapping between the two programs, which is something that
-relational verification work seeks to answer. Thus DAC will be considered further after the idea of product programs is exhausted.
-
-% How is vmv related to relational verification?
-
-% Standard taint analysis will under/overapproximate the taint, so we may not be able to use it to directly prove our safety property on the target program.
-Thus, we can leverage the approach verification modulo versions to prove relational verification.
-
-% Relational verification aims to solve this problem by proving relational properties between two programs.
-However, the naive relational verification approach of verifying each program individually is infeasible given arbitrarily difficult-to-reason-about functions even if those functions do not change between versions of the program.
-One approach is to underapproximate the taint on one program and overapproximate the taint on another program to verify the set equality of taint sources, but this still falls prey to the approximation of taint analysis given arbitrarily complex functions.
-Also, current tooling does not widely support underapproximating taint analysis.
 

### 2018-12-13 paper/paper.tex
figures
Notes and typos

@@
     Additionally, we run taint analysis on each program to get outputs $\Gamma_1$ and $\Gamma_2$, which we use in the next step.
-    \item \textbf{Program syncronization:} Using our environment $\Gamma = \Gamma_1 \cup \Gamma_2$, our instrumented programs $P'_1$ and $P'_2$ from 
+    \item \textbf{Program synchronization:} Using our environment $\Gamma = \Gamma_1 \cup \Gamma_2$, our instrumented programs $P'_1$ and $P'_2$ from 
     the previous step, we construct the product program $P'_1 \times P'_2$. This converts the $2$-safety problem to one of $1$-safety.
@@
 
-	% The next step involves collecting static information about the program that will help us apply our inference rules. Namely, we will populate the environment $\Gamma$ such that
+    % TODO: talk about what makes us say some functions are preserving and some are not
 
-  %   \item Replace sources with labelled constants
-  %   \item For values that are tainted by more than one source (for example $S_1 + S_2$) replace with one of two uninterpreted functions over the sources: \begin{enumerate}
-		% \item $\textsf{preserving}(s_1, s_2, \ldots, s_n)$. Ex: $+, \oplus$
-		% \item $\textsf{non-preserving}(s_1, s_2, \ldots, s_n)$. Ex: $<<, >>$
-  %   \end{enumerate}
-  %   \item Perform taint analysis on sources to generate environment $\Gamma$ which marks statements involving tainted variables.
 \textbf{Instrumentation:} Given two program versions $P_1$ and $P_2$, our sources $\Sigma$, and a taint analysis algorithm $\mathsf{Taint}_{S,L}$

### 2018-12-14 paper/paper.tex
paper stuff

@@
 \begin{figure}
+\caption{}
 \label{fig:semrules}
-\caption{}
 \[
@@
 \begin{itemize}
-    \item $\Gamma \gets \mathsf{Taint}_{S,L}(\phi, T, \Sigma, Z, P)$: Takes as input an initial assignment $\phi$ of some taint sets $T = \{\tau_i\}$ to 
+    \item $\Gamma \gets \mathsf{Taint}_{S,L}(\phi, T, \Sigma, P)$: Takes as input an initial assignment $\phi$ of some taint sets $T = \{\tau_i\}$ to 
     a set of sources $\Sigma = \{\sigma_i\}$, and an input program $P$. Outputs an 
@@
 \begin{figure}
+    \caption{Taint analysis overapproximation}
     \label{fig:prog1}
-    \caption{Taint analysis overapproximation}
     \begin{center}
@@
 \begin{figure}
+    \caption{Taint analysis unsoundness}
     \label{fig:prog2}
-    \caption{Taint analysis unsoundness}
     \begin{center}
@@
 
-Description of differential taint analysis.
+In this section, we introduce the notion of \emph{differential taint analysis}.
+We consider two programs $P_1$ and $P_2$ with shared
+source set $\Sigma$, taint set $T$, and sink set $Z$, and source-taint mapping $\phi$. 
+A differential taint analysis algorithm
+should accept only if for every program trace $\sigma_1$ and $\sigma_2$ of $P_1$
+and $P_2$, respectively, on the same input $\vec{x}$ the taint set for all sinks
+$\zeta_1 \in Z$ in $P_1$ are equal to the taint set for all sinks $\zeta_2 \in Z$ in $P_2$.
+This notion is meant to capture the idea that if $P_1$ has all correct program traces,
+$P_2$ will as well.
 
@@
 Ultimately, our goal is to make assertions about the taintsets of program variables at the sinks.
-But first, we will individually instrument each version of the software project.
+But first we will individually instrument each version of the software project.
 
@@
 	\begin{enumerate}
-		\item $\textsf{preserving}(s_1, s_2, \ldots, s_n)$ if the operation preserves entropy (+, XOR, etc.)
-		\item $\textsf{non-preserving}(s_1, s_2, \ldots, s_n)$ if the operation does not preserve entropy (<<, >>, etc.)
+		\item $\textsf{preserving}(s_1, s_2, \ldots, s_n)$ if the operation preserves entropy 
+		\item $\textsf{non-preserving}(s_1, s_2, \ldots, s_n)$ if the operation does not preserve entropy 
 	\end{enumerate}
@@
 
-    % TODO: talk about what makes us say some functions are preserving and some are not
+Formally, we propose the following algorithm:
 
-\textbf{Instrumentation:} Given two program versions $P_1$ and $P_2$, our sources $\Sigma$, and a taint analysis algorithm $\mathsf{Taint}_{S,L}$
-     we reduce the taint analysis problem to one of $1$-safety.
-    We do this for each $P_k$ by assigning constants $c$ to all sources $\sigma \in \Sigma$, and then where we would propagate taint using the rules $S$ for language $L$
-    in statements $s_i$ which is a function of program variables $v_i$ and only of ``entropy preserving" operations (described below) 
-    we replace with an uninterpreted function $f(v_1, \ldots, v_j)$. For statements $s'_i$ which are a
-    function of any tainted program variables $w_j$ and contain any ``non-entropy preserving operations'' (any operations that are not entropy preserving), we replace
-    with an uninterpreted function $g(w_1, \ldots, w_j)$. We output these programs as $P'_1$ and $P'_2$.
+\begin{itemize}
+    \item $(\Gamma, P') \gets \mathsf{Instrument}(P, \mathsf{Taint}_{S,L}, \Sigma, \Delta)$: Given a program $P$, a taint analysis subroutine $\mathsf{Taint}_{S,L}$, a set
+    of sources $\Sigma$, and a set of entropy-preserving operations $\Delta \subseteq \mathsf{Ops}$, we create a new program $P'$ in the following way. First,
+    we compute $\Gamma \gets \mathsf{Taint}_{S,L}(\phi, \{\top\}, \Sigma, P)$ where $\phi$ is the membership function of $\Sigma$. Then:
+    \begin{itemize}
+        \item For each source $\sigma \in \Sigma$, replace with a labelled constant $c$.
+        \item For all variables $x_i$ which are an assignment of a sequence of program variables $v_i$ and combined by operations $\odot \in \mathsf{Ops}$,
+        if any $v_i$ is tainted and all operations are in $\Delta$ then assign $x_i$ to an uninterpreted function $\textsf{preserving}(v_1, \ldots, v_j)$. If
+        any $v_i$ is tainted but one operations is not in $\Delta$, then assign $x_i$ to an uninterpreted function $\textsf{non-preserving}(v_1, \ldots, v_j)$.
+        Otherwise, do nothing.
+    \end{itemize}
+    Output $\Gamma$ and $P'$.
+\end{itemize}
 
-Description of instrumentation algorithm.
+We only consider operations known to preserve entropy such as addition (for example: over some modulus, like $\oplus$, or adding signed fixed-length integers, or arbitrary-precision) for inclusion in set $\Delta$.
+Other operations, like bit-shifts, must be excluded from $\Delta$ to preserve security.
 
@@
 \begin{figure}
+    \caption{Basic Inference Rules}
     \label{fig:infrules1}
-    \caption{Basic Inference Rules}
     \[
@@
 \begin{figure}
+    \caption{Conditional and Loop Inference Rules}
     \label{fig:infrules2}
-    \caption{Conditional and Loop Inference Rules}
     \[
@@
 \begin{figure}
+    \caption{Augmented Inference Rules}
     \label{fig:auginfrules1}
-    \caption{Augmented Inference Rules}
     \[
@@
 \begin{figure}
+    \caption{Augmented Inference Rules}
     \label{fig:auginfrules2}
-    \caption{Augmented Inference Rules}
     \[

### 2018-12-14 paper/paper.tex
Added examples

--- 2018-12-13 paper/paper.tex
@@
 Also, because the variables in the two versions of the program have been renamed to be disjoint, we can introduce an inference rule for the commutativity of two statements.
+See Figure \ref{fig:infrules1}.
 
 \begin{figure}
+    \caption{Basic Inference Rules}
     \label{fig:infrules1}
-    \caption{Basic Inference Rules}
     \[
@@
 The benefit of this transformation, however, is that the predicate abstraction tool has more information about the trace taken with which it can reason about product program.
+See \ref{fig:infrules2}.
 
@@
 To maintain the equivalence with the sequential composition, we also run the remaining iterations in the leftover loop after the other loop's condition no longer holds.
+See \ref{fig:infrules2}.
 
 \begin{figure}
+    \caption{Conditional and Loop Inference Rules}
     \label{fig:infrules2}
-    \caption{Conditional and Loop Inference Rules}
     \[
@@
 
+To demonstrate the application of these inference rules we will work an example. Consider the following two programs: Algorithm \ref{alg:p1} and \ref{alg:p2}.
+
+\begin{algorithm}
+  \caption{$P1$}
+  \label{alg:p1}
+  \begin{algorithmic}[1]
+    \If {$p$}
+        \State $x_1 \gets 1$
+    \Else
+        \State $x_1 \gets 2$
+    \EndIf
+  \end{algorithmic}
+\end{algorithm}
+
+and
+
+\begin{algorithm}
+  \caption{$P1$}
+  \label{alg:p2}
+  \begin{algorithmic}[1]
+    \If {$p$}
+        \State $x_1 \gets 2$
+    \Else
+        \State $x_1 \gets 1$
+    \EndIf
+  \end{algorithmic}
+\end{algorithm}
+
+These programs just switch the bodies of the if and else branches (so they are clearly not equivalent).
+
+Sequentially composed, we could imagine passing the following program to CPAChecker: Algorithm \ref{alg:seq}.
+
+\begin{algorithm}
+  \caption{$P1; P2$}
+  \label{alg:seq}
+  \begin{algorithmic}[1]
+    \If {$p$}
+        \State $x_1 \gets 1$
+    \Else
+        \State $x_1 \gets 2$
+    \EndIf
+    \If {$p$}
+        \State $x_1 \gets 2$
+    \Else
+        \State $x_1 \gets 1$
+    \EndIf
+    \State \textsf{assert}$(x_1 == x_2)$
+  \end{algorithmic}
+\end{algorithm}
+
+But instead, we could synchronize the if statements with the if-statement inference rule to yield: Algorithm \ref{alg:cross}.
+
+\begin{algorithm}
+  \caption{$P1{} \cross P2$}
+  \label{alg:cross}
+  \begin{algorithmic}[1]
+    \If {$p$}
+        \State $x_1 \gets 1$
+        \If {$p$}
+            \State $x_1 \gets 2$
+        \Else
+            \State $x_1 \gets 1$
+        \EndIf
+    \Else
+        \State $x_1 \gets 2$
+        \If {$p$}
+            \State $x_1 \gets 2$
+        \Else
+            \State $x_1 \gets 1$
+        \EndIf
+    \EndIf
+    \State \textsf{assert}$(x_1 == x_2)$
+  \end{algorithmic}
+\end{algorithm}
+
+Clearly Algorithm \ref{alg:cross} is twice as long because we replicated $P2$ into both branches of the if-statement.
+However this is a tradeoff we may be willing to make in order to ease the burden on our predicate abstraction tool as compared to Algorithm \ref{alg:seq}.
+
 \subsection{Taint-augmented Inference Rules}
@@
 We introduce an optimization inference rule that does not lead to any exponential blowup for statements that are not entailed by $\Gamma$.
+See \ref{fig:auginfrules1}.
 
 \begin{figure}
+    \caption{Augmented Inference Rules}
     \label{fig:auginfrules1}
-    \caption{Augmented Inference Rules}
     \[
@@
 
-Then, we can make the corresponsding changes to our naive inference rules to recognize the $\Gamma$ environment we've introduced.
-This gives us the following set of augmented inference rules: % TODO: reference to figure
+Then, we can make the corresponding changes to our naive inference rules to recognize the $\Gamma$ environment we've introduced.
+This gives us the following set of augmented inference rules: See Figure \ref{fig:auginfrules2}.
 
 \begin{figure}
+    \caption{Augmented Inference Rules}
     \label{fig:auginfrules2}
-    \caption{Augmented Inference Rules}
     \[

### 2018-12-14 paper/paper.tex

--- 2018-12-14 paper/paper.tex
@@
 Also, because the variables in the two versions of the program have been renamed to be disjoint, we can introduce an inference rule for the commutativity of two statements.
+See Figure \ref{fig:infrules1}.
 
@@
 To maintain the equivalence with the sequential composition, we also run the remaining iterations in the leftover loop after the other loop's condition no longer holds.
+See \ref{fig:infrules2}.
 
@@
 We introduce an optimization inference rule that does not lead to any exponential blowup for statements that are not entailed by $\Gamma$.
+See \ref{fig:auginfrules1}.
 
@@
 
-Then, we can make the corresponsding changes to our naive inference rules to recognize the $\Gamma$ environment we've introduced.
-This gives us the following set of augmented inference rules: % TODO: reference to figure
+Then, we can make the corresponding changes to our naive inference rules to recognize the $\Gamma$ environment we've introduced.
+This gives us the following set of augmented inference rules: See Figure \ref{fig:auginfrules2}.
 

### 2018-12-14 paper/paper.tex
refs and stuff

@@
 \subsection{Related Work}
+
+Verification modulo version \cite{logozzo2014verification}.
 
@@
 
-% \section{Future Work}
+\section{Future Work}
 
-% Verify the following conjectures. Prove that a heuristic-based approach is no less powerful than a full product program approach.
-% Run this tool on projects in the wild to determine how widespread entropy-misuse bugs are among software projects ``in the wild.''
+Verify the following conjectures. Prove that a heuristic-based approach is no less powerful than a full product program approach.
+Run this tool on projects in the wild to determine how widespread entropy-misuse bugs are among software projects ``in the wild.''
+
+{\normalsize \bibliographystyle{acm}
+\bibliography{refs}}
 
@@
 
-\section{Links}
+% \section{Links}
 
-\begin{enumerate}
-	\item Debian/OpenSSL Bug 
-		\begin{enumerate}
-			\item \url{https://www.schneier.com/blog/archives/2008/05/random_number_b.html}
-			\item \url{https://research.swtch.com/openssl}
-			\item \url{https://freedom-to-tinker.com/2013/09/20/software-transparency-debian-openssl-bug/}
-			\item \url{https://www.cs.umd.edu/class/fall2017/cmsc818O/papers/private-keys-public.pdf}
-		\end{enumerate}
-	\item Data flow
-		\begin{enumerate}
-			\item \url{https://en.wikipedia.org/wiki/Data-flow_analysis}
-			\item \url{https://www.seas.harvard.edu/courses/cs252/2011sp/slides/Lec02-Dataflow.pdf}
-		\end{enumerate}
-	\item Static Program Analysis
-		\begin{enumerate}
-            \item \url{https://cs.au.dk/~amoeller/spa/spa.pdf}
-            \item \url{https://ieeexplore.ieee.org/stamp/stamp.jsp?arnumber=6859783}
-		\end{enumerate}
-	\item Relational Verification:
-		\begin{enumerate}
-			\item \url{https://dl.acm.org/citation.cfm?id=2021319}
-			\item \url{https://ac.els-cdn.com/S235222081630044X/1-s2.0-S235222081630044X-main.pdf?_tid=076a0492-9cee-4995-9710-bcb3c64b98e0&acdnat=1539815890_178849b4f14af3751e9acb03b238db4d}
-			\item \url{https://www.microsoft.com/en-us/research/publication/differential-assertion-checking/}
-			\item \url{https://www.microsoft.com/en-us/research/wp-content/uploads/2014/06/paper-1.pdf}
-			\item \url{https://www.cs.utexas.edu/~isil/pldi16-chl.pdf}
-		\end{enumerate}
-	\item Projects to analyze 
-		\begin{enumerate}
-			\item OpenPGP
-			\item BouncyCastle
-			\item OpenSSL
-			\item GnuPGP
-			\item F\# SSL project with proof of correctness
-			\item NQSBTLS
-			\item Amazon's s2n (signal to noise)
-		\end{enumerate}
-\end{enumerate}
-
-{\normalsize \bibliographystyle{acm}
-\bibliography{refs}}
+% \begin{enumerate}
+% 	\item Debian/OpenSSL Bug 
+% 		\begin{enumerate}
+% 			\item \url{https://www.schneier.com/blog/archives/2008/05/random_number_b.html}
+% 			\item \url{https://research.swtch.com/openssl}
+% 			\item \url{https://freedom-to-tinker.com/2013/09/20/software-transparency-debian-openssl-bug/}
+% 			\item \url{https://www.cs.umd.edu/class/fall2017/cmsc818O/papers/private-keys-public.pdf}
+% 		\end{enumerate}
+% 	\item Data flow
+% 		\begin{enumerate}
+% 			\item \url{https://en.wikipedia.org/wiki/Data-flow_analysis}
+% 			\item \url{https://www.seas.harvard.edu/courses/cs252/2011sp/slides/Lec02-Dataflow.pdf}
+% 		\end{enumerate}
+% 	\item Static Program Analysis
+% 		\begin{enumerate}
+%             \item \url{https://cs.au.dk/~amoeller/spa/spa.pdf}
+%             \item \url{https://ieeexplore.ieee.org/stamp/stamp.jsp?arnumber=6859783}
+% 		\end{enumerate}
+% 	\item Relational Verification:
+% 		\begin{enumerate}
+% 			\item \url{https://dl.acm.org/citation.cfm?id=2021319}
+% 			\item \url{https://ac.els-cdn.com/S235222081630044X/1-s2.0-S235222081630044X-main.pdf?_tid=076a0492-9cee-4995-9710-bcb3c64b98e0&acdnat=1539815890_178849b4f14af3751e9acb03b238db4d}
+% 			\item \url{https://www.microsoft.com/en-us/research/publication/differential-assertion-checking/}
+% 			\item \url{https://www.microsoft.com/en-us/research/wp-content/uploads/2014/06/paper-1.pdf}
+% 			\item \url{https://www.cs.utexas.edu/~isil/pldi16-chl.pdf}
+% 		\end{enumerate}
+% 	\item Projects to analyze 
+% 		\begin{enumerate}
+% 			\item OpenPGP
+% 			\item BouncyCastle
+% 			\item OpenSSL
+% 			\item GnuPGP
+% 			\item F\# SSL project with proof of correctness
+% 			\item NQSBTLS
+% 			\item Amazon's s2n (signal to noise)
+% 		\end{enumerate}
+% \end{enumerate}
 

### 2018-12-14 paper/paper.tex
Related work

@@
 
-Verification modulo version \cite{logozzo2014verification}.
+\subsubsection{Differential assertion checking}
 
-Differential assertion checking, verification modulo versions, relational verification using product programs, secure information flow as a safety problem
+This paper generates relative specifications between multiple versions of the program, much in the same way we generate the relative specification between the taintsets of variables that flow into sinks.
+However, the differential assertion checking paper focuses on managing the number of warnings emitted by static analysis programs by using the correctness of previous versions to silence errors.
+In contrast, our approach aims to surface an entirely new class of warnings.
+See \cite{lahiri2013differential}.
+
+\subsubsection{Secure information flow as a safety problem}
+This paper explores the notion of applying type information of low-security types and high-security types to guide the inference rules in the construction of product programs.
+It paints a traditional 2-safety property of secure information flow as a 1-safety property by constructing the product program.
+Rather than using multiple versions of the same program like our approach, this paper explores the self-composition of one program, which is especially helpful for synchronizing statements when reducing from a 2-safety property.
+See  \cite{terauchi2005secure}.
+
+\subsubsection{Relational verification using product programs}
+
+This paper explores the inference rules needed to construct product programs for relational verification.
+Primarily, the inference rules synchronize loops in an effort to simplify the process of problems like inferring loop invariants to optimize the loops.
+Our work builds on this work by introducing the notion of using taint analysis environment to guide the inference rules.
+See \cite{barthe2011relational}.
+
+\subsubsection{Verification modulo versions}
+
+This paper leverages the multiple versions of a program to tackle the problem of spurious warnings in a sound way.
+It speaks to a larger trend of the hypothesis that multiple versions of a program generated in the software development lifecycle today can be capitalized by the static analysis community.
+See \cite{logozzo2014verification}.
 

### 2018-12-14 paper/paper.tex

--- 2018-12-14 paper/paper.tex
@@
 \begin{abstract}
- Test abstract
+    In this work we examine the problem of entropy misuse vulnerabilities in
+    security-critical software. Our work approaches the problem from an
+    ``ounce of prevention is worth a pound of cure'' perspective; by introducing
+    and applying a novel static analysis technique \emph{differential taint analysis},
+    we propose a method for catching these bugs. Our approach
+    leverages modern software development's
+     common practice of maintaining software version history to 
+     provide augmented heuristics to standard static analysis tools.
 \end{abstract}
@@
 \subsection{Predicate Abstraction}
+\label{sec:predabstraction}
 
@@
 \begin{itemize}
-    \item $b \gets \mathsf{Verify}_{L}(\pi, P)$: Takes as input a $1$-safety property $\pi$ and a program $P$ and outputs an accepting bit $\top$ if
-    $P$ satisfies the safety property, and otherwise outputs a rejecting bit $\bot$ or fails to terminate otherwise.
+    \item $b \gets \mathsf{Verify}_{L}(\Pi, P)$: Takes as input a set $\Pi$ of $1$-safety properties and a program $P$ and outputs an accepting bit $\top$ if
+    $P$ satisfies all $\pi \in \Pi$, and otherwise outputs a rejecting bit $\bot$ or fails to terminate otherwise.
 \end{itemize}
@@
 
-\subsection{Naive Product Program Inference Rules}
+\subsection{Basic Product Program Inference Rules}
+\label{sec:basicpprules}
 
@@
 \subsection{Taint-augmented Inference Rules}
+\label{sec:augpprules}
 
@@
 See \ref{fig:auginfrules1}.
+
+Formally, we define the following algorithm:
+
+\begin{itemize}
+    \item $P_1 \times P_2 \gets \mathsf{Prod}_R(P_1, P_2, \Gamma)$: Given two programs $P_1$ and $P_2$, a taint environment (mapping of program statements) to 
+    taint values), and a set of inference rules $R$ we output the product program $P_1 \times P_2$.
+\end{itemize}
 
@@
 
-Description of algorithm using the two subroutines above.
+Let $\Delta$ be some set of entropy-preserving operations in our language $L$ with some set of 
+natural taint propagation rules $S$.
+Let $R$ be our set of inference rules described in Sections \ref{sec:basicpprules} and \ref{sec:augpprules}.
+Using the subroutines we described above and the verification method in Section
+\ref{sec:predabstraction}, we have the following algorithm:
+
+\begin{itemize}
+    \item $b \gets \mathsf{DifferentialTaintAnalysis}_\Delta(P_1, P_2, \Sigma, Z)$: We take 
+    as input two programs $P_1$ and $P_2$, a set of sources $\Sigma$, and a set of 
+    sinks $Z$. We then compute:
+    \begin{itemize}
+        \item $(\Gamma_1, P_1') \gets \mathsf{Instrument}(P_1, \mathsf{Taint}_{S,L}, \Sigma, \Delta)$
+        \item $(\Gamma_2, P_2') \gets \mathsf{Instrument}(P_2, \mathsf{Taint}_{S,L}, \Sigma, \Delta)$
+        \item $P'_1 \times P'_2 \gets \mathsf{Prod}_R(P'_1, P'_2, \Gamma_1 \cup \Gamma_2)$
+    \end{itemize}
+    At each synchronized sink pair $(\zeta_1, \zeta_2) \in Z$ in $P'_1 \times P'_2$, add an assertion of equality 
+    $(\zeta_1 = \zeta_2)$ to a set of safety properties $\Pi$. Then compute $b \gets \mathsf{Verify}_{L}(\Pi, P'_1 \times P'_2)$
+    and output $b$.
+\end{itemize}
 
@@
 
-Verify the following conjectures. Prove that a heuristic-based approach is no less powerful than a full product program approach.
-Run this tool on projects in the wild to determine how widespread entropy-misuse bugs are among software projects ``in the wild.''
+In continuing this work, we would like to pursue a couple main lines of inquiries:
+
+\begin{itemize}
+    \item We wish to see how much our taint-augmented inference rules aid the verifier, from an experimental perspective on large projects such as OpenSSL.
+    \item We wish to prove that our taint-augmented inferences rules (or some superset of them) are ``just as useful'' as the basic rules (that is,
+    our verifier does not lose any information from using the scoped-down rules).
+    \item We want a formal reduction between a notion of differential taint analysis and the approach described here.
+    \item Create a push-button software solution for large projects, suitable for insertion in a continuous integration development cycle.
+    \item Take an at-large survey of open-source software for entropy misuse vulnerabilities.
+\end{itemize}
 

### 2018-12-14 paper/paper.tex
Evaluation

@@
 \subsection{Putting it all together}
+\label{sec:all-together}
 
@@
 
-% TODO
+To evaluate the algorithm laid out in Section \ref{sec:all-together}, we utilized an off-the-shelf state-of-the-art predicate abstraction tool called CPAChecker \cite{beyer2011cpachecker}.
+This tool was selected because it was the winner in SV-COMP'18, the Competition on Software Verification.
+
+We demonstrated the motivating examples by writing a simple C program for each one and introducing entropy bugs to demonstrate possible issues.
+We selected C because most entropy bugs like the OpenSSL bug and the FreeBSD bug were implemented in systems level code written in C.
+Then, we instrumented these programs by replacing the sources and propagating them appropriately.
+After applying the inference rules to the composition of the original program and the buggy program, we ended up with a synchronized product program.
+We also applied the inference rules on the self-composition of the program to demonstrate version changes that don't introduce bugs that should be verified.
+The necessary assertions were inserted at the sinks after the product program was constructed.
+
+By plugging these product programs into CPAChecker we were able to get the desired verification results.
+CPAChecker was able to successfully verify the correct changes and successfully fail on the incorrect changes.
+
+Note that our motivating examples were small and simple.
+Thus, the augmented inference rules evaluated to the same result as the naive inference rules.
+Our evaluation code snippets were correctly solved by CPAChecker in every variation of our inference rules: the sequential composition, the synchronized composition, and the augmented composition.
+Further evaluation is needed to confirm or reject our conjectures about the scalability of the various approaches in the design space and to explore the impact of our augmented inferene rules.
 

### 2018-12-14 paper/paper.tex

--- 2018-12-14 paper/paper.tex
@@
     we propose a method for catching these bugs. Our approach
-    leverages modern software development's
-     common practice of maintaining software version history to 
-     provide augmented heuristics to standard static analysis tools.
+    leverages the widespread practice of maintaining software version history for
+    modern software projects in order to
+     provide augmented heuristics to standard static analysis tools to prove the correctness of
+     subsequent versions of programs relative to the correctness of an original version.
+     Our contributions include the introduction of differential taint analysis
+     as a novel approach to this problem domain, as well as a construction to solve
+     the differential taint analysis problem using a reduction to standard static
+     analysis techniques.
 \end{abstract}
@@
 
-One common misuse of cryptography is the misuse of entropy. Without proper random inputs, many cryptographic algorithms
-are vulnerable to basic forms of cryptanalysis. Some cryptographic schemes, such as DSA, can even disclose long-term secrets
-such as the signing key when the random per-message input is low entropy, made public, or nonunique.
+One common misuse of cryptography is the misuse of entropy. Without properly random inputs, many cryptographic schemes
+are vulnerable to basic forms of cryptanalysis. Some cryptographic algorithms, such as DSA, can even disclose long-term secrets
+such as the signing key when the random per-message input is low entropy, made public, or nonunique. Such frailty implies that
+proper entropy use is a necessary requirement for secure systems.
 
-We plan to use data dependency tools to determine how entropic inputs in a given program are used by various cryptographic algorithms.
-That will allow us to identify if and when entropy is too low or is misused.
-We plan to produce this as a code integration tool for developers to use as part of a compiler toolchain.
-Our tool will seek to generate an error report from source code using notions of taint analysis.
-
-For this project we consider codebases that maintain version history, in the hopes that we can employ static analysis techniques across program versions to infer more information about how entropy is being used;
-our primary goal is to identify bugs that are introduced to existing codebases.
-This may prove to be especially helpful since naive taint analysis can provide mere overapproximations of data dependency for a given program and may generate many false alarms for programmers in a professional development setting.
-
-\subsection{Related Work}
-
-\subsubsection{Differential assertion checking}
-
-This paper generates relative specifications between multiple versions of the program, much in the same way we generate the relative specification between the taintsets of variables that flow into sinks.
-However, the differential assertion checking paper focuses on managing the number of warnings emitted by static analysis programs by using the correctness of previous versions to silence errors.
-In contrast, our approach aims to surface an entirely new class of warnings.
-See \cite{lahiri2013differential}.
-
-\subsubsection{Secure information flow as a safety problem}
-This paper explores the notion of applying type information of low-security types and high-security types to guide the inference rules in the construction of product programs.
-It paints a traditional 2-safety property of secure information flow as a 1-safety property by constructing the product program.
-Rather than using multiple versions of the same program like our approach, this paper explores the self-composition of one program, which is especially helpful for synchronizing statements when reducing from a 2-safety property.
-See  \cite{terauchi2005secure}.
-
-\subsubsection{Relational verification using product programs}
-
-This paper explores the inference rules needed to construct product programs for relational verification.
-Primarily, the inference rules synchronize loops in an effort to simplify the process of problems like inferring loop invariants to optimize the loops.
-Our work builds on this work by introducing the notion of using taint analysis environment to guide the inference rules.
-See \cite{barthe2011relational}.
-
-\subsubsection{Verification modulo versions}
-
-This paper leverages the multiple versions of a program to tackle the problem of spurious warnings in a sound way.
-It speaks to a larger trend of the hypothesis that multiple versions of a program generated in the software development lifecycle today can be capitalized by the static analysis community.
-See \cite{logozzo2014verification}.
+Detecting this class of bug often requires an examination of the source code; 
+issues where outputs from cryptographic algorithms are completely deterministic
 
@@
 the absence of errors in a single program trace \cite{alpern1987recognizing}, a $k$-safety property of a program requires the absence of erroneous interactions between $k$
-traces of the same program \cite{sousa2016cartesian}. A program property such as symmetry is a $2$-safety property.
+traces of the same program or single traces of $k$ different programs \cite{sousa2016cartesian}. A program property such as symmetry is a $2$-safety property. In the context of this
+work we only consider $2$-safety properties.
 
@@
 
-% TODO
+\section{Related Work}
+
+In this section we explore a few other results in the area of differential or relational program verification.
+
+\subsection{Differential assertion checking}
+
+This paper generates relative specifications between multiple versions of the program, much in the same way we generate the relative specification between the taintsets of variables that flow into sinks.
+However, the differential assertion checking paper focuses on managing the number of warnings emitted by static analysis programs by using the correctness of previous versions to silence errors.
+In contrast, our approach aims to surface an entirely new class of warnings.
+See \cite{lahiri2013differential}.
+
+\subsection{Secure information flow as a safety problem}
+This paper explores the notion of applying type information of low-security types and high-security types to guide the inference rules in the construction of product programs.
+It paints a traditional 2-safety property of secure information flow as a 1-safety property by constructing the product program.
+Rather than using multiple versions of the same program like our approach, this paper explores the self-composition of one program, which is especially helpful for synchronizing statements when reducing from a 2-safety property.
+See  \cite{terauchi2005secure}.
+
+\subsection{Relational verification using product programs}
+
+This paper explores the inference rules needed to construct product programs for relational verification.
+Primarily, the inference rules synchronize loops in an effort to simplify the process of problems like inferring loop invariants to optimize the loops.
+Our work builds on this work by introducing the notion of using taint analysis environment to guide the inference rules.
+See \cite{barthe2011relational}.
+
+\subsection{Verification modulo versions}
+
+This paper leverages the multiple versions of a program to tackle the problem of spurious warnings in a sound way.
+It speaks to a larger trend of the hypothesis that multiple versions of a program generated in the software development lifecycle today can be capitalized by the static analysis community.
+See \cite{logozzo2014verification}.
+
+\section{Future Work}
+
+In continuing this work, we would like to pursue a couple of different lines of inquiries:
+
+\begin{itemize}
+    \item How much do our taint-augmented inference rules aid the verifier, from an experimental perspective on large projects such as OpenSSL?
+    \item We wish to prove that our taint-augmented inferences rules (or some superset of them) are ``just as useful'' as the basic rules (that is,
+    our verifier does not lose any information from using the scoped-down rules).
+    \item We want a formal reduction between a notion of differential taint analysis and the approach described here.
+    \item Create a push-button software solution for large projects, suitable for insertion in a continuous integration development cycle.
+    \item Take an at-large survey of open-source software for entropy misuse vulnerabilities.
+    \item What other security-related applications does differential taint analysis have?
+\end{itemize}
 
@@
 found in the Debian OpenSSL and FreeBSD projects.
-
-\section{Future Work}
-
-In continuing this work, we would like to pursue a couple main lines of inquiries:
-
-\begin{itemize}
-    \item We wish to see how much our taint-augmented inference rules aid the verifier, from an experimental perspective on large projects such as OpenSSL.
-    \item We wish to prove that our taint-augmented inferences rules (or some superset of them) are ``just as useful'' as the basic rules (that is,
-    our verifier does not lose any information from using the scoped-down rules).
-    \item We want a formal reduction between a notion of differential taint analysis and the approach described here.
-    \item Create a push-button software solution for large projects, suitable for insertion in a continuous integration development cycle.
-    \item Take an at-large survey of open-source software for entropy misuse vulnerabilities.
-\end{itemize}
 

### 2018-12-14 paper/paper.tex

@@
 \label{sec:eval}
+
+To evaluate the algorithm laid out in Section \ref{sec:all-together}, we utilized an off-the-shelf state-of-the-art predicate abstraction tool called CPAChecker \cite{beyer2011cpachecker}.
+This tool was selected because it was the winner in SV-COMP'18, the Competition on Software Verification.
+
+We demonstrated the motivating examples by writing a simple C program for each one and introducing entropy bugs to demonstrate possible issues.
+We selected C because most entropy bugs like the OpenSSL bug and the FreeBSD bug were implemented in systems level code written in C.
+Then, we instrumented these programs by replacing the sources and propagating them appropriately.
+After applying the inference rules to the composition of the original program and the buggy program, we ended up with a synchronized product program.
+We also applied the inference rules on the self-composition of the program to demonstrate version changes that don't introduce bugs that should be verified.
+The necessary assertions were inserted at the sinks after the product program was constructed.
+
+By plugging these product programs into CPAChecker we were able to get the desired verification results.
+CPAChecker was able to successfully verify the correct changes and successfully fail on the incorrect changes.
+
+Note that our motivating examples were small and simple.
+Thus, the augmented inference rules evaluated to the same result as the naive inference rules.
+Our evaluation code snippets were correctly solved by CPAChecker in every variation of our inference rules: the sequential composition, the synchronized composition, and the augmented composition.
+Further evaluation is needed to confirm or reject our conjectures about the scalability of the various approaches in the design space and to explore the impact of our augmented inferene rules.
 

### 2018-12-14 paper/paper.tex

@@
 
-In this work we present a new static analysis technique to aid in the prevention of entropy-misuse security vulnerabilities, such as those
-found in the Debian OpenSSL and FreeBSD projects.
+In this work we explored solutions to the entropy vulnerabilities in large software projects.
+This was motivated by real-world entropy propagation failures found in the Debian OpenSSL project from 2008 and the FreeBSD project from 2016.
+Our primary contribution was proposing a novel technique for the relational verification of new versions of a program modulo the audited program.
+
+Our proposed solution blends taint analysis with relational verification to control the tractability of the safety analysis task.
+The algorithm proceeds in three primary steps: instrumenting the project, constructing the product program, and inserting assertions/verifying with off-the-shelf static analysis tools.
+These steps implement two reductions: the instrumentation step converts the original taint analysis problem into a 2-safety property, and the product program construction converts the 2-safety property into a 1-safety property that can be statically verified after insertions are injected in the final step.
+These constructed product programs were evaluated with an off-the-shelf static analysis tool called CPAChecker.
+
+Three options are explored for the construction of the product program to sufficiently consider the tradeoff between ease of reasoning for the predicate abstraction tool and the potential for exponential blowup in the size of the product program.
+A hybrid is suggested that leverages the domain specific knowledge of entropy-tainted variables to find a suitable middle ground.
+The initial results presented in this paper represent a promising opportunity for future work.
 

### 2018-12-14 paper/paper.tex
work

--- 2018-12-14 paper/paper.tex
@@
 such as the signing key when the random per-message input is low entropy, made public, or nonunique. Such frailty implies that
-proper entropy use is a necessary requirement for secure systems.
+proper entropy use is a necessary requirement for secure systems. These bugs have appeared
+in widely-used software projects. We briefly discuss two historical examples.
+
+\subsection{Debian OpenSSL}
+
+\cite{schneier2008}
+
+\subsection{FreeBSD}
+
+\cite{hovav2016}
+
+\subsection{Approaches}
 
 Detecting this class of bug often requires an examination of the source code; 
-issues where outputs from cryptographic algorithms are completely deterministic
+bugs that lead to completely deterministic outputs
+ from cryptographic algorithms can often be caught easily during the manual
+ testing of code before it is released. Issues which persist past release
+ typically result from low-entropy inputs rather than no-entropy inputs.
+ As such, these bugs can be difficult to discover and track down. One solution
+ to this problem, manual auditing, does work (historically, a manual
+ inspection of source code is how these vulnerabilities are disclosed and fixed).
+ However effective, manual audits can be an expensive solution, and altogether 
+ infeasible
+ for many projects to do often, especially those reliant on community donations and gratis
+ developer work. On the other hand, static analysis tools have their own issues:
+ due to classic hardness results, static analysis algorithms must rely on heuristic
+ methods and often cannot provide the sufficient guarantees to developers, even
+ if they admit soundness for all of their results.
+
+ \subsection{Contributions}
 

### 2018-12-14 paper/paper.tex
more life -drake -michael scott

@@
     security-critical software. Our work approaches the problem from an
-    ``ounce of prevention is worth a pound of cure'' perspective; by introducing
-    and applying a novel static analysis technique \emph{differential taint analysis},
+    ``ounce of prevention is worth a pound of cure'' perspective (we do not
+    contribute new disclosures of this bug); by introducing
+    and applying a novel static analysis technique called \emph{differential taint analysis},
     we propose a method for catching these bugs. Our approach
@@
 \subsection{Debian OpenSSL}
+\label{sec:openssl}
 
-\cite{schneier2008}
+The Debian project maintains its own version of OpenSSL. In 2008 a Debian developer
+pushed a patch to the Debian OpenSSL project which commented out a few lines of code
+marked by Valgrind, a tool
+commonly used to identify memory leak bugs. One particular line 
+in OpenSSL's random number generator code was one of these lines \cite{schneier2008}:
+
+\begin{center}
+\begin{verbatim}
+MD_Update(&m,buf,j);
+\end{verbatim}
+\end{center}
+
+Unfortunately, this line seeded OpenSSL's
+RNG with kernel randomness. As a result, the only entropy that the RNG received
+was from the process's ID, which is capped at a value of 32,768 on Linux.
+This implies that there are only 32,768 possible outputs for the RNG, which
+is used by software to generate cryptographic keys. So all keys generated by
+this flawed RNG could be easily enumerated by any adversary.
 
 \subsection{FreeBSD}
+\label{sec:freebsd}
 
-\cite{hovav2016}
+An entropy failure bug was also found in the FreeBSD project in 2016 \cite{hovav2016}.
+This particular vulnerability was a result of the kernel relying on a weak random number
+generator (a simple linear congruential generator, well-known to be
+vulnerable to basic cryptanalysis) seeded with the current time during bootup. 
+After boot the kernel code was supposed to switch to a
+source of high-entropy (presumably at this point in time an entropy pool would be
+sufficiently full). However, this switch to the secure adaptor was never made, and the kernel continued 
+to rely on the weak RNG throughout the lifetime of the system. This bug is particularly
+alarming because many userspace applications, including popular security products like 
+OpenSSL, rely on the kernel
+to generate high-entropy randomness on their behalf. In fact, kernel randomness is
+the de-facto recommended source of entropy in the security community.
 
@@
  testing of code before it is released. Issues which persist past release
- typically result from low-entropy inputs rather than no-entropy inputs.
+ typically result from low-entropy inputs rather than no-entropy inputs, such as
+ the historical examples discussed in sections \ref{openssl} and \ref{freebsd}.
  As such, these bugs can be difficult to discover and track down. One solution
@@
 
+ In this work, we propose an approach which marries the best of both worlds.
+ We introduce a new static analysis technique we call \emph{differential taint analysis}
+ which seeks to prove relative or differential correctness between programs (or program versions).
+ This allows software projects to only pay for a single audit of their code, and use
+ our methods to prove that future updates are relatively secure to this original version.
+ Intuitively, proving relational properties between similar programs should be easier;
+ in fact, as in other works \cite{barthe2011relational, lahiri2013differential, logozzo2014verification, sousa2016cartesian}
+ we reduce the problem to verifying safety properties of a \emph{single} program, and use domain-specific
+ properties (taint analysis) to reduce the burden of the verifier.
+
+ \subsection{Outline}
+
+ In section \ref{sec:prelim} we describe the static analysis background necessary
+ to understand our results. We then define differential
+ taint analysis and the details of our methods of reducing it to standard
+ analysis techniques in section \ref{sec:approach}. Finally, we discuss a 
+ simple evaluation of our methods in section \ref{sec:eval}.
+
 \section{Preliminaries}
+\label{sec:prelim}
 
@@
 \section{Our Approach}
+\label{sec:approach}
 
@@
 	These uninterpreted functions are inferred to be pure functions over their variables, so they capture the notion of the taintset at that point.
-This allows us to assert their equality in the program we pass to CPAChecker.
+This allows us to assert their equality in the program we pass to CPAchecker.
 
@@
 
-Although these inference rules are semantically equivalent to the sequential composition, and clearly provide benefit to CPAChecker, this proving strength comes at a cost.
+Although these inference rules are semantically equivalent to the sequential composition, and clearly provide benefit to CPAchecker, this proving strength comes at a cost.
 Namely, the if statements and the while loops lead to an exponential blowup in the size of the product program being reasoned about as a result of the copied statements.
@@
 
-and
-
 \begin{algorithm}
@@
 
-Sequentially composed, we could imagine passing the following program to CPAChecker: Algorithm \ref{alg:seq}.
+Sequentially composed, we could imagine passing the following program to CPAchecker: Algorithm \ref{alg:seq}.
 
@@
 
-To evaluate the algorithm laid out in Section \ref{sec:all-together}, we utilized an off-the-shelf state-of-the-art predicate abstraction tool called CPAChecker \cite{beyer2011cpachecker}.
+To evaluate the algorithm laid out in Section \ref{sec:all-together}, we utilized an off-the-shelf state-of-the-art predicate abstraction tool called CPAchecker \cite{beyer2011cpachecker}.
 This tool was selected because it was the winner in SV-COMP'18, the Competition on Software Verification.
@@
 
-By plugging these product programs into CPAChecker we were able to get the desired verification results.
-CPAChecker was able to successfully verify the correct changes and successfully fail on the incorrect changes.
+By plugging these product programs into CPAchecker we were able to get the desired verification results.
+CPAchecker was able to successfully verify the correct changes and successfully fail on the incorrect changes.
 
@@
 Thus, the augmented inference rules evaluated to the same result as the naive inference rules.
-Our evaluation code snippets were correctly solved by CPAChecker in every variation of our inference rules: the sequential composition, the synchronized composition, and the augmented composition.
+Our evaluation code snippets were correctly solved by CPAchecker in every variation of our inference rules: the sequential composition, the synchronized composition, and the augmented composition.
 Further evaluation is needed to confirm or reject our conjectures about the scalability of the various approaches in the design space and to explore the impact of our augmented inferene rules.
@@
 These steps implement two reductions: the instrumentation step converts the original taint analysis problem into a 2-safety property, and the product program construction converts the 2-safety property into a 1-safety property that can be statically verified after insertions are injected in the final step.
-These constructed product programs were evaluated with an off-the-shelf static analysis tool called CPAChecker.
+These constructed product programs were evaluated with an off-the-shelf static analysis tool called CPAchecker.
 
